---
title: "Remove in-port VMS records"
output: html_document
---

Date: 12/18/2018


The purpose of this script is to flag all "in-port" VMS records, using a buffer zone of 3km / 1.5km. In-port records with an average speed of < 1 are marked for removal. Only those records marked for removal before and after the fishing trip were actually removed from the filtered output data. 
Original version (12/17/2018) focused only on trips targeting Dungeness crab.

*Version 2: performs filtering using the calculate average speed, rather than the reported average speed.*


<br>
```{r "setup", include=FALSE}
knitr::opts_knit$set(root.dir = "D:/VMS-repo") 
```
<br>
<br>

### Prep for running code
Clear workspace
```{r}
rm(list=ls())
```
<br>

Install packages
```{r include=FALSE}
#install.packages("foreign")
#install.packages("dplyr")
#install.packages("tidyr")
#install.packages("lubridate") #https://rstudio-pubs-static.s3.amazonaws.com/28038_1bcb9aa80ca84f27ace07d612872861a.html

library(foreign)
library(dplyr)
library(tidyr)
library(lubridate) 
library(ggplot2)
library(maps)
library(rgdal)
library(rgeos)
```
<br>

Create Objects / Data Frames
```{r}
## directory where Blake's processed VMS data is stored as .dbf files
origVMS_dir = "ProcessVMS/R_Output/match/unfiltered/cleaned/"

## directory where the filtered VMS data will be written out
filterVMS_dir = "ProcessVMS/R_Output/match/unfiltered/cleaned/"

## file with port coordinates, including directory
coords_file = "Input_Data/port_coords_fromBlake_edited.csv"

## calendar years
years <- c(2011)

## width of buffer circle (in meters). applies to all ports but those that require smaller buffers
r = 3000

## port codes for any ports that require reduced buffer sizes. Default is only COS
ports_lowbuffer <- c("COS")

## width of buffer circle (in meters) for COS
r2 = 1500

## cutoff value for speed (m/s) -- 1m/s ~ 2 knots
speed_cutoff <- 0.75

## do you want to filter out data points that are potentially on shore (NGDC > 0m)?
bathy_filter=FALSE

## do you want a ton of plots for quality control?
plot.me = FALSE
```
<br>

### Read in the data

First, the full vms data set
```{r rawdat}
vms <- read.csv(paste0(origVMS_dir,"VMS_Outputs_wTARGET_10d_lookback_", years[1], "_speed_cleaned.csv"))
dim(vms)
head(vms)
```
<br>

Then, the coordinates for the ports
```{r}
portdat <- read.csv(coords_file)
if(any(is.na(portdat$lat))==TRUE){
  message("WARNING: some ports are missing coordinates. ports were removed from data frame.")
  portdat <- filter(portdat, !is.na(lat))
}
port_coords <- portdat
port_coords_reg <- portdat %>% filter(!(port_code %in% ports_lowbuffer))
port_coords_sm <- portdat %>% filter(port_code %in% ports_lowbuffer)
```
<br>

### Perform optional bathymetry filter
```{r}
if(bathy_filter == TRUE){
  vms <- vms %>%
    filter(NGDC_M <= 0)
  vms_removed <- vms %>%
    filter(NGDC_M > 0)
  vms_removed$westcoastdate <- ymd_hms(vms_removed$westcoastdate, tz="America/Los_Angeles")
  for(i in seq(1, length(years))){
    tmp_dat <- filter(vms_removed, year(westcoastdate) == years[i])
    write.csv(tmp_dat, paste0(filterVMS_dir, "VMS_Outputs_wTARGET_10d_lookback_", years[i], "_speed_cleaned_bathy_REMOVED.csv", row.names=FALSE))
  }
}
```
<br>


### Prep the port coordinates
*Note - to assign a project, `proj4string` requires coordinates in xy (lng/lat) order*
```{r}
coordinates(port_coords) <- c("Lon", "Lat") 
proj4string(port_coords) <- CRS("+init=epsg:4326") # WGS 84
port_coords_geo <- spTransform(port_coords, CRSobj = CRS("+init=epsg:32610"))

coordinates(port_coords_reg) <- c("Lon", "Lat") 
proj4string(port_coords_reg) <- CRS("+init=epsg:4326") # WGS 84
port_coords_reg_geo <- spTransform(port_coords_reg, CRSobj = CRS("+init=epsg:32610"))

coordinates(port_coords_sm) <- c("Lon", "Lat") 
proj4string(port_coords_sm) <- CRS("+init=epsg:4326") # WGS 84
port_coords_sm_geo <- spTransform(port_coords_sm, CRSobj = CRS("+init=epsg:32610"))
```
<br>

### Create buffer
```{r buffer}
buffer_polygons_reg <- gBuffer(port_coords_reg_geo, byid=TRUE, width=r)
buffer_polygons_sm <- gBuffer(port_coords_sm_geo, byid=TRUE, width=r2)
buffer_polygons  <- rbind(buffer_polygons_reg, buffer_polygons_sm)
```


Get correct input format / coordinate system for ggplot
```{r}
ports_df <- data.frame(port_coords)
buffer_polygons_LL <- spTransform(buffer_polygons, CRSobj = CRS("+init=epsg:4326"))
plot(buffer_polygons_LL)
```

Basemap
```{r}
data(stateMapEnv)
states_df <- map_data("state") %>%
  filter(region == "california" | region=="oregon" | region=="washington")
```

Plot coastline
```{r}
ggplot() +
  geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
  geom_point(data=ports_df, aes(x=Lon, y=Lat), col="red") +
  coord_cartesian(xlim=c(-127, -112), ylim=c(32, 49)) +
  geom_polygon(data=buffer_polygons_LL, aes(x = long, y = lat, group = group), col="blue") +
  geom_text(data=ports_df, aes(x=Lon, y=Lat, label=port_code),size=2, nudge_x=-0.51) 
```

Zoom in to a bay area and a coast area (*this is a crappy background map, so if ports are on land, that's ok.*)
```{r}
cutcoast <- list(c(-124, -125, 47.5, 48.5), #upper WA
                 c(-123.5, -124.5, 46.7, 47.7), #lower WA p1
                 c(-123.5, -124.5, 45.7, 46.7), #lower WA / upper OR [WLB, NHL]
                 c(-123.5, -124.5,45.6, 44.6), # OR [TLL, NEW]
                 c(-123.5, -124.5,44.5, 43.5), # OR [WLD, WIN]
                 c(-124, -125, 43.5, 42.25), # OR [COS,GLD]
                 c(-123.75, -124.75, 42.2, 40.5), #OR to CA [BRK, FLN]
                 c(-123, -124, 40, 38.5), # CA [BRG, ARE]
                 c(-122, -123.5, 38.5, 36.6), # CA [BDG, CRZ]
                 c(-120.5, -122, 37, 35), # CA [MOS, AVL]
                 c(-117, -120, 35, 32.5)) # CA[SB, OCN]
```
```{r}
for(i in seq(1,11)){
  myplot <- ggplot() +
    geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
    geom_polygon(data=buffer_polygons_LL, aes(x = long, y = lat, group = group), col="blue") +
    geom_point(data=ports_df, aes(x=Lon, y=Lat), col="red", size=2) +
    coord_cartesian(xlim=c(cutcoast[[i]][1], cutcoast[[i]][2]), ylim=c(cutcoast[[i]][3], cutcoast[[i]][4])) +
    geom_text(data=ports_df, aes(x=Lon, y=Lat, label=port_code),size=4, nudge_x=-0.07)
  print(myplot)
}
```

Zoom in to a bay area and a coast area WITH VMS POINTS. (*this is a crappy background map, so if ports are on land, that's ok.*)
```{r}
if(plot.me){
  for(i in seq(1:11)){
    myplot <- ggplot() +
      geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
      geom_polygon(data=buffer_polygons_LL, aes(x = long, y = lat, group = group), col="blue") +
      geom_point(data=ports_df, aes(x=Lon, y=Lat), col="red", size=2) +
      geom_point(data=vms, aes(x=LONGITUDE, y=LATITUDE), col="green", alpha=0.5) +
      coord_cartesian(xlim=c(cutcoast[[i]][1], cutcoast[[i]][2]), ylim=c(cutcoast[[i]][3], cutcoast[[i]][4]))
    print(myplot)
  }
} else{
  for(i in seq(from=1, to=9, by=3)){
    myplot <- ggplot() +
      geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
      geom_polygon(data=buffer_polygons_LL, aes(x = long, y = lat, group = group), col="blue") +
      geom_point(data=ports_df, aes(x=Lon, y=Lat), col="red", size=2) +
      geom_point(data=vms, aes(x=LONGITUDE, y=LATITUDE), col="green", alpha=0.5) +
      coord_cartesian(xlim=c(cutcoast[[i]][1], cutcoast[[i]][2]), ylim=c(cutcoast[[i]][3], cutcoast[[i]][4]))
    print(myplot)
  }
}

```
<br>


### Flag data points
https://www.nceas.ucsb.edu/scicomp/usecases/point-in-polygon 

First, let's just make sure that our polygon objects are set up correctly. Do this by verifying that all port coordinates are in the correct port polygon buffer. 
```{r inport_flag}
port_coords_geo$test_port <- over(port_coords_geo, buffer_polygons)$port_code
port_coords_geo$port_code[port_coords_geo$port_code != port_coords_geo$test_port]
port_coords_geo$test_port[port_coords_geo$port_code != port_coords_geo$test_port]
```
Some Santa Barbara and LA area ports may overlap buffers >= 3km.
<br>


Assign projection to VMS points
```{r}
vms_sp <- vms
coordinates(vms_sp) <- c("LONGITUDE", "LATITUDE") 
proj4string(vms_sp) <- CRS("+init=epsg:4326") # WGS 84
vms_geo <- spTransform(vms_sp, CRSobj = CRS("+init=epsg:32610"))
```
<br>


Identify data points within buffer zones
```{r}
in.port <- over(vms_geo, buffer_polygons)$port_code
vms <- mutate(vms, in_port = in.port)
```
<br>


Zoom in to a bay area and a coast area WITH VMS POINTS. (*this is a crappy background map, so if ports are on land, that's ok.*)
```{r}
if(plot.me){
  for(i in seq(1,11)){
    myplot <- ggplot() +
      geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
      geom_polygon(data=buffer_polygons_LL, aes(x = long, y = lat, group = group), col="blue") +
      geom_point(data=ports_df, aes(x=Lon, y=Lat), col="red", size=2) +
      geom_point(data=vms, aes(x=LONGITUDE, y=LATITUDE, col=in_port), alpha=0.5) +
      coord_cartesian(xlim=c(cutcoast[[i]][1], cutcoast[[i]][2]), ylim=c(cutcoast[[i]][3], cutcoast[[i]][4])) +
      geom_text(data=ports_df, aes(x=Lon, y=Lat, label=port_code),size=4, nudge_x=0.07) 
    print(myplot)
  }
}
```
<br>

What is the avg speed when vessels are marked as in port? 
```{r}
inport_dat <- mutate(vms, in_port_binary = ifelse(is.na(in_port), 0,1))
ggplot(data=filter(inport_dat, avg_speed_ms < 50), aes(x=avg_speed_ms)) +
  geom_histogram() +
  facet_wrap(~in_port_binary) +
  xlab("Average Speed (Calculated)") +
  ylab("Number of Records")
```
<br>

Mark "remove" if the record is in port *and* the avg speed is < 1 m/s.
```{r remove_flag}
vms <- mutate(vms, to_rm = ifelse(!is.na(in_port) & avg_speed_ms < speed_cutoff, "remove", "retain"))
cat("Proportion of VMS records removed: ", sum(vms$to_rm == "remove") / dim(vms)[1], "\n")
cat("There were ", dim(vms)[1]/sum(vms$to_rm == "retain"), "x VMS records in the unfiltered data than will be retained.\n")
cat("Proportion of VMS records marked in-port removed: ", sum(vms$to_rm == "remove") / sum(!is.na(vms$in_port)))
```
<br>


### Filtering data

I want to remove all but the last in-port record before the fishing trip, and the first in-port record after the fishing trip. For now, I am leaving in all mid-trip "remove" points. 

To do this, I first need to get the date / time of the first and last "retain" record for each trip. Then, I can split each trip as pre-first "retain" and post-last "retain". Finally, sort by date/timestamp and take the top n=1 VMS record per trip. Add these data back into the output data frame.

Get the `westcoastdate` of the first/last record to retain
```{r}
vms$westcoastdate <- ymd_hms(vms$westcoastdate)
first_last_retain <- vms %>%
  filter(to_rm == "retain") %>%
  group_by(Rec_ID) %>%
  summarise(first_retain = min(westcoastdate, na.rm=TRUE), last_retain = max(westcoastdate, na.rm=TRUE))
vms_ext <- left_join(vms, first_last_retain, by="Rec_ID")
```
<br>

Split "remove" records into pre-trip and post-trip data
```{r split_remove}
vms_ext$westcoastdate <- ymd_hms(vms_ext$westcoastdate)
pretrip_inport <- vms_ext %>%
  filter(westcoastdate < first_retain) %>%
  filter(to_rm == "remove")
posttrip_inport <- vms_ext %>%
  filter(westcoastdate > last_retain) %>%
  filter(to_rm == "remove")
```
<br>

Sort pre-trip by desc(westcoastdate), grab the top record per trip. Sort post-trip by westcoastdate, grab the top record per trip.
```{r}
pretrip_lastvms <- pretrip_inport %>%
  group_by(Rec_ID) %>%
  top_n(n=1, wt=westcoastdate)
posttrip_firstvms <- posttrip_inport %>%
  group_by(Rec_ID) %>%
  top_n(n=1, wt=desc(westcoastdate))
head(pretrip_lastvms %>% dplyr::select(Rec_ID, westcoastdate, to_rm, first_retain))
head(posttrip_firstvms %>% dplyr::select(Rec_ID, westcoastdate, to_rm, last_retain))
```
<br>

Remove records pre-first_retain and post-last_retain
```{r filter}
vms_ext_filtered <- vms_ext %>%
  filter(westcoastdate > first_retain & westcoastdate < last_retain)
vms_ext_removed <- vms_ext %>%
  filter(westcoastdate < first_retain | westcoastdate > last_retain)
```
<br>

Integrate last pre-trip and first post-trip back into the data set. 
```{r add_one_inport}
colnames(vms_ext_filtered) == colnames(pretrip_lastvms)
colnames(vms_ext_filtered) == colnames(posttrip_firstvms)

vms_out_ext <- vms_ext_filtered %>%
  bind_rows(pretrip_lastvms) %>%
  bind_rows(posttrip_firstvms)
vms_out_ext$westcoastdate <- ymd_hms(vms_out_ext$westcoastdate)
vms_out_ext <- vms_out_ext %>%
  arrange(Rec_ID, westcoastdate)
vms_out <- dplyr::select(vms_out_ext, c(-first_retain, -last_retain))
```
<br>

Write out filtered data set, with a separate file for each year
```{r write_out}
if(bathy_filter == TRUE){
  for(y in years){
    tmp_output <- filter(vms_out, year(westcoastdate) == y)
    tmp_output$westcoastdate <- as.character(tmp_output$westcoastdate)
    write.csv(x=tmp_output, paste0(filterVMS_dir, "VMS_Outputs_wTARGET_10d_lookback_", y, "_speed_cleaned_bathy_offshore.csv"), row.names=FALSE)
    tmp_filteredout <- filter(vms_ext_removed, year(westcoastdate) == y)
    tmp_filteredout$westcoastdate <- as.character(tmp_filteredout$westcoastdate)
    write.csv(x=tmp_filteredout, paste0(filterVMS_dir, "VMS_Outputs_wTARGET_10d_lookback_", y, "_speed_cleaned_bathy_inport_removed.csv"), row.names=FALSE)
  }
} else{
  for(y in years){
    tmp_output <- filter(vms_out, year(westcoastdate) == y)
    tmp_output$westcoastdate <- as.character(tmp_output$westcoastdate)
    write.csv(x=tmp_output, paste0(filterVMS_dir, "VMS_Outputs_wTARGET_10d_lookback_", y, "_speed_cleaned_offshore.csv"), row.names=FALSE)
    tmp_filteredout <- filter(vms_ext_removed, year(westcoastdate) == y)
    tmp_filteredout$westcoastdate <- as.character(tmp_filteredout$westcoastdate)
    write.csv(x=tmp_filteredout, paste0(filterVMS_dir, "VMS_Outputs_wTARGET_10d_lookback_", y, "_speed_cleaned_inport_removed.csv"), row.names=FALSE)
  }
}
```
<br>
<br>



_________________________________
### View flagged points

#### In port, removed
```{r plot_remove}
rm_vms <- filter(vms, to_rm=="remove")
if(plot.me){
  for(i in seq(1,11)){
    myplot<- ggplot() +
      geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
      geom_polygon(data=buffer_polygons_LL, aes(x = long, y = lat, group = group), col="blue") +
      geom_point(data=ports_df, aes(x=Lon, y=Lat), col="red", size=2) +
      geom_point(data=rm_vms, aes(x=LONGITUDE, y=LATITUDE, col=avg_speed_ms), alpha=0.5) +
      coord_cartesian(xlim=c(cutcoast[[i]][1], cutcoast[[i]][2]), ylim=c(cutcoast[[i]][3], cutcoast[[i]][4])) +
      geom_text(data=ports_df, aes(x=Lon, y=Lat, label=port_code),size=4, nudge_x=0.07) +
      ggtitle("Removed VMS Records")
    print(myplot)
  }
} else{
  for(i in seq(from=1,to=9, by=3)){
    myplot<- ggplot() +
      geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
      geom_polygon(data=buffer_polygons_LL, aes(x = long, y = lat, group = group), col="blue") +
      geom_point(data=ports_df, aes(x=Lon, y=Lat), col="red", size=2) +
      geom_point(data=rm_vms, aes(x=LONGITUDE, y=LATITUDE, col=avg_speed_ms), alpha=0.5) +
      coord_cartesian(xlim=c(cutcoast[[i]][1], cutcoast[[i]][2]), ylim=c(cutcoast[[i]][3], cutcoast[[i]][4])) +
      geom_text(data=ports_df, aes(x=Lon, y=Lat, label=port_code),size=4, nudge_x=0.07) +
      ggtitle("Removed VMS Records")
    print(myplot)
  }
}
```
<br>

#### In port, retained
```{r plot_retain}
retain_vms <- filter(vms, to_rm=="retain" & !is.na(in_port))
if(plot.me){
  for(i in seq(1,11)){
    myplot<- ggplot() +
      geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
      geom_polygon(data=buffer_polygons_LL, aes(x = long, y = lat, group = group), col="blue") +
      geom_point(data=ports_df, aes(x=Lon, y=Lat), col="red", size=2) +
      geom_point(data=retain_vms, aes(x=LONGITUDE, y=LATITUDE, col=avg_speed_ms), alpha=0.5) +
      coord_cartesian(xlim=c(cutcoast[[i]][1], cutcoast[[i]][2]), ylim=c(cutcoast[[i]][3], cutcoast[[i]][4])) +
      geom_text(data=ports_df, aes(x=Lon, y=Lat, label=port_code),size=4, nudge_x=0.07) +
      ggtitle("Retained In-Port VMS Records")
    print(myplot)
  }
}  else{
  for(i in seq(from=1,to=9, by=3)){
    myplot<- ggplot() +
      geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
      geom_polygon(data=buffer_polygons_LL, aes(x = long, y = lat, group = group), col="blue") +
      geom_point(data=ports_df, aes(x=Lon, y=Lat), col="red", size=2) +
      geom_point(data=rm_vms, aes(x=LONGITUDE, y=LATITUDE, col=avg_speed_ms), alpha=0.5) +
      coord_cartesian(xlim=c(cutcoast[[i]][1], cutcoast[[i]][2]), ylim=c(cutcoast[[i]][3], cutcoast[[i]][4])) +
      geom_text(data=ports_df, aes(x=Lon, y=Lat, label=port_code),size=4, nudge_x=0.07) +
      ggtitle("Removed VMS Records")
    print(myplot)
  }
}
```



### Check effects of flagging

#### Have we removed > 99% of records from any trips?
```{r plot_premove}
premoved <- vms %>%
  group_by(Port_Of_Landing, Rec_ID) %>%
  summarise(p.remove = sum(to_rm=="remove") / n())
cat("Removed more than 99% of records from ", sum(premoved$p.remove > 0.99) / length(premoved$p.remove) * 100, "% (" , sum(premoved$p.remove > 0.99) , "total) of trips. ")
ggplot(premoved, aes(x=p.remove)) +
  geom_histogram() +
  ggtitle("Across All Ports") 
if(plot.me){
  for(p in premoved$Port_Of_Landing){
    myplot <- ggplot(data=filter(premoved, Port_Of_Landing == p), aes(x=p.remove)) +
      geom_histogram() +
      ggtitle(paste0("Port: ", p)) +
      xlab("Percent VMS records removed per trip") +
      ggtitle(p)
    print(myplot)
  }
}  else{
  for(i in seq(from=1,to=9, by=3)){
    myplot<- ggplot() +
      geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
      geom_polygon(data=buffer_polygons_LL, aes(x = long, y = lat, group = group), col="blue") +
      geom_point(data=ports_df, aes(x=Lon, y=Lat), col="red", size=2) +
      geom_point(data=rm_vms, aes(x=LONGITUDE, y=LATITUDE, col=avg_speed_ms), alpha=0.5) +
      coord_cartesian(xlim=c(cutcoast[[i]][1], cutcoast[[i]][2]), ylim=c(cutcoast[[i]][3], cutcoast[[i]][4])) +
      geom_text(data=ports_df, aes(x=Lon, y=Lat, label=port_code),size=4, nudge_x=0.07) +
      ggtitle("Removed VMS Records")
    print(myplot)
  }
}
```
<br>



#### General viz across ports
Paths of 2 trips from each port: original data set
```{r plot_2trips}
if(plot.me){
  for(p in unique(vms$Port_Of_Landing)){
    tmp_port <- dplyr::filter(vms, Port_Of_Landing == p)
    if(length(unique(tmp_port$Rec_ID)) > 1){
      for(i in seq(1,2)){
        tmp_dat <- filter(tmp_port, Rec_ID == unique(tmp_port$Rec_ID)[i])
        label_dat <- rbind(tmp_dat[1,], tmp_dat[length(tmp_dat$Rec_ID),])
        myplot <- ggplot(tmp_dat) +
          geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
          geom_point(aes(x=LONGITUDE, y=LATITUDE, col=to_rm), size = 4, alpha = 0.5) + 
          geom_path(aes(x=LONGITUDE, y=LATITUDE), col="dodgerblue4", size = 0.5, alpha = 0.5) +
          geom_text(data=label_dat, aes(y = LATITUDE,x=LONGITUDE, label=c("S", "E"))) +
          geom_point(data=filter(ports_df, as.character(port_code) == as.character(unique(tmp_dat$Port_Of_Landing))), aes(x=Lon, y=Lat), col="red", size=2) +
          #scale_color_continuous() +
          ggtitle(paste0("Trip ID:", i, "Port:", tmp_dat$Port_Of_Landing[1])) +
          coord_cartesian(xlim=c(min(tmp_dat$LONGITUDE) - 0.1, max(tmp_dat$LONGITUDE) + 0.1), ylim=c(min(tmp_dat$LATITUDE) - 0.1, max(tmp_dat$LATITUDE) + 0.1))
        png(paste0("ProcessVMS/R_Output/quality/buffer_zones/",years[1], "/", p, "_Rec", unique(tmp_dat$Rec_ID), ".png"))
        print(myplot)
        dev.off()
      }
    }
    else{
      tmp_dat <- tmp_port
      label_dat <- rbind(tmp_dat[1,], tmp_dat[length(tmp_dat$Rec_ID),])
      myplot <- ggplot(tmp_dat) +
        geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
        geom_point(aes(x=LONGITUDE, y=LATITUDE, col=to_rm), size = 4, alpha = 0.5) + 
        geom_path(aes(x=LONGITUDE, y=LATITUDE), col="dodgerblue4", size = 0.5, alpha = 0.5) +
        geom_text(data=label_dat, aes(y = LATITUDE,x=LONGITUDE, label=c("S", "E"))) +
        geom_point(data=filter(ports_df, as.character(port_code) == as.character(unique(tmp_dat$Port_Of_Landing))), aes(x=Lon, y=Lat), col="red", size=2) +
        #scale_color_continuous() +
        ggtitle(paste0("Trip ID:", i, "Port:", tmp_dat$Port_Of_Landing[1])) +
        coord_cartesian(xlim=c(min(tmp_dat$LONGITUDE) - 0.1, max(tmp_dat$LONGITUDE) + 0.1), ylim=c(min(tmp_dat$LATITUDE) - 0.1, max(tmp_dat$LATITUDE) + 0.1))
      png(paste0("ProcessVMS/R_Output/quality/buffer_zones/",years[1], "/", p, "_Rec", unique(tmp_dat$Rec_ID), ".png"))
      print(myplot)
      dev.off()
    }
  }
}
```




Paths of 2 trips from each port: newly filtered data set
```{r plot_2trips_new}
if(plot.me){
  for(p in unique(vms_out$Port_Of_Landing)){
    tmp_port <- dplyr::filter(vms_out, Port_Of_Landing == p)
    if(length(unique(tmp_port$Rec_ID)) > 1){
      for(i in seq(1,2)){
        tmp_dat <- filter(tmp_port, Rec_ID == unique(tmp_port$Rec_ID)[i])
        label_dat <- rbind(tmp_dat[1,], tmp_dat[length(tmp_dat$Rec_ID),])
        myplot <- ggplot(tmp_dat) +
          geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
          geom_point(aes(x=LONGITUDE, y=LATITUDE, col=to_rm), size = 4, alpha = 0.5) + 
          geom_path(aes(x=LONGITUDE, y=LATITUDE), col="dodgerblue4", size = 0.5, alpha = 0.5) +
          geom_text(data=label_dat, aes(y = LATITUDE,x=LONGITUDE, label=c("S", "E"))) +
          geom_point(data=filter(ports_df, as.character(port_code) == as.character(unique(tmp_dat$Port_Of_Landing))), aes(x=Lon, y=Lat), col="red", size=2) +
          #scale_color_continuous() +
          ggtitle(paste0("Trip ID:", i, "Port:", tmp_dat$Port_Of_Landing[1])) +
          coord_cartesian(xlim=c(min(tmp_dat$LONGITUDE) - 0.1, max(tmp_dat$LONGITUDE) + 0.1), ylim=c(min(tmp_dat$LATITUDE) - 0.1, max(tmp_dat$LATITUDE) + 0.1))
        png(paste0("ProcessVMS/R_Output/quality/buffer_zones/", years[1], "/filter_", p, "_Rec", unique(tmp_dat$Rec_ID), ".png"))
        print(myplot)
        dev.off()
      }
    }
    else{
      tmp_dat <- tmp_port
      label_dat <- rbind(tmp_dat[1,], tmp_dat[length(tmp_dat$Rec_ID),])
      myplot <- ggplot(tmp_dat) +
        geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
        geom_point(aes(x=LONGITUDE, y=LATITUDE, col=to_rm), size = 4, alpha = 0.5) + 
        geom_path(aes(x=LONGITUDE, y=LATITUDE), col="dodgerblue4", size = 0.5, alpha = 0.5) +
        geom_text(data=label_dat, aes(y = LATITUDE,x=LONGITUDE, label=c("S", "E"))) +
        geom_point(data=filter(ports_df, as.character(port_code) == as.character(unique(tmp_dat$Port_Of_Landing))), aes(x=Lon, y=Lat), col="red", size=2) +
        #scale_color_continuous() +
        ggtitle(paste0("Trip ID:", i, "Port:", tmp_dat$Port_Of_Landing[1])) +
        coord_cartesian(xlim=c(min(tmp_dat$LONGITUDE) - 0.1, max(tmp_dat$LONGITUDE) + 0.1), ylim=c(min(tmp_dat$LATITUDE) - 0.1, max(tmp_dat$LATITUDE) + 0.1))
      png(paste0("ProcessVMS/R_Output/quality/buffer_zones/", years[1], "/filter_", p, "_Rec", unique(tmp_dat$Rec_ID), ".png"))
      print(myplot)
      dev.off()
    }
  }
}
```








