---
title: "Mid-Trip Returns Assessment"
output: html_document
---

I was left with more questions on the apparent mid-trip returns occurring in the data set. This script takes a look at the phenomena on both a broader and individual trip scale. The assignment of in-port points to specific port locations was used for some of the questions below.




<br>
```{r "setup", include=FALSE}
knitr::opts_knit$set(root.dir = "D:/VMS-repo") 
```
<br>
<br>

### Prep for running code
Clear workspace
```{r}
rm(list=ls())
```
<br>

Install packages
```{r include=FALSE}
#install.packages("foreign")
#install.packages("dplyr")
#install.packages("tidyr")
#install.packages("lubridate") #https://rstudio-pubs-static.s3.amazonaws.com/28038_1bcb9aa80ca84f27ace07d612872861a.html

library(foreign)
library(dplyr)
library(tidyr)
library(lubridate) 
library(ggplot2)
library(maps)
library(rgdal)
library(rgeos)
```
<br>

Create Objects / Data Frames
```{r}
## directory where Blake's processed VMS data is stored as .dbf files
origVMS_dir = "ProcessVMS/R_Output/match/unfiltered/cleaned/"

## directory where the filtered VMS data will be written out
filterVMS_dir = "ProcessVMS/R_Output/match/unfiltered/cleaned/"

## file with port coordinates, including directory
coords_file = "Input_Data/port_coords_fromBlake.csv"

## calendar years
years <- c(2010)

## width of buffer circle (in meters)
r = 3000
```
<br>

### Read in the data

First, the full vms data set
```{r}
vms <- read.csv(paste0(origVMS_dir,"VMS_Outputs_wTARGET_10d_lookback_2010_cleaned.csv"))
dim(vms)
head(vms)
dcrb_vms <- filter(vms, TARGET_max == "DCRB")
```
<br>

Then, the coordinates for the ports
```{r}
portdat <- read.csv(coords_file)
if(any(is.na(portdat$lat))==TRUE){
  message("WARNING: some ports are missing coordinates. ports were removed from data frame.")
  portdat <- filter(portdat, !is.na(lat))
}
port_coords <- portdat
```


### Prep the port coordinates
*Note - to assign a project, `proj4string` requires coordinates in xy (lng/lat) order*
```{r}
coordinates(port_coords) <- c("Lon", "Lat") 
proj4string(port_coords) <- CRS("+init=epsg:4326") # WGS 84
port_coords_geo <- spTransform(port_coords, CRSobj = CRS("+init=epsg:32610"))
```
<br>

### Create buffer
```{r}
buffer_polygons <- gBuffer(port_coords_geo, byid=TRUE, width=r)
```
<br>

Get correct input format / coordinate system for ggplot
```{r}
ports_df <- data.frame(port_coords)
buffer_polygons_LL <- spTransform(buffer_polygons, CRSobj = CRS("+init=epsg:4326"))
plot(buffer_polygons_LL)
```
<br>

Basemap
```{r}
data(stateMapEnv)
states_df <- map_data("state") %>%
  filter(region == "california" | region=="oregon" | region=="washington")
```

Plot coastline
```{r}
ggplot() +
  geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
  geom_point(data=ports_df, aes(x=Lon, y=Lat), col="red") +
  coord_cartesian(xlim=c(-127, -112), ylim=c(32, 49)) +
  geom_polygon(data=buffer_polygons_LL, aes(x = long, y = lat, group = group), col="blue") +
  geom_text(data=ports_df, aes(x=Lon, y=Lat, label=port_code),size=2, nudge_x=-0.51) 
```
<br>

Assign projection to VMS points
```{r}
dcrb_vms_sp <- dcrb_vms
coordinates(dcrb_vms_sp) <- c("LONGITUDE", "LATITUDE") 
proj4string(dcrb_vms_sp) <- CRS("+init=epsg:4326") # WGS 84
dcrb_vms_geo <- spTransform(dcrb_vms_sp, CRSobj = CRS("+init=epsg:32610"))
```
<br>


### Identify data points within buffer zones
```{r}
in.port <- over(dcrb_vms_geo, buffer_polygons)$port_code
dcrb_vms <- mutate(dcrb_vms, in_port = in.port)
```
<br>


Mark "remove" if the record is in port *and* the avg speed is < 0.05.
```{r}
dcrb_vms <- mutate(dcrb_vms, to_rm = ifelse(!is.na(in_port) & AVG_SPEED < 1, "remove", "retain"))
cat("Proportion of VMS records removed: ", sum(dcrb_vms$to_rm == "remove") / dim(dcrb_vms)[1], "\n")
cat("There were ", dim(dcrb_vms)[1]/sum(dcrb_vms$to_rm == "retain"), "x VMS records in the unfiltered data than will be retained.\n")
cat("Proportion of VMS records marked in-port removed: ", sum(dcrb_vms$to_rm == "remove") / sum(!is.na(dcrb_vms$in_port)))
```
<br>

### Remove in-port points at start / end of trip

First, I need to get the time of the first and last "retain" record in each trip.
```{r}
dcrb_vms$westcoastdate <- ymd_hms(dcrb_vms$westcoastdate)
time_retain <- dcrb_vms %>%
  filter(to_rm == "retain") %>%
  group_by(Rec_ID) %>%
  summarise(first_retain = min(westcoastdate), last_retain = max(westcoastdate))
dcrb_vms_ext <- left_join(dcrb_vms, time_retain, by="Rec_ID")
```
<br>

Then I can remove all rows marked "to remove" that are earlier than the first "retain" record, or later than the last "retain" record.
```{r}
dcrb_vms_part_rm <- dcrb_vms_ext %>%
  filter(westcoastdate > first_retain & westcoastdate < last_retain)
```
<br>


Write this out:
```{r}
write.csv(dcrb_vms_part_rm, "ProcessVMS/R_Output/match/unfiltered/cleaned/DCRB_VMS_Outputs_wTARGET_10d_lookback_2010_tripedge_inport_rm.csv",
          row.names=FALSE)
```
<br>


### Assessment

#### What proportion of trips still have data points marked "to remove", overall and by port group?
```{r}
midtrip_removals <- dcrb_vms_part_rm %>%
  group_by(Rec_ID, Port_Of_Landing) %>%
  summarise(midtrip = ifelse("remove" %in% to_rm, "y", "n"))

count_midtrip_removals <- midtrip_removals %>%
  group_by(Port_Of_Landing) %>%
  summarise(ntrips = sum(midtrip == "y"), ptrips = ntrips / n())
```
<br>

```{r}
ggplot(count_midtrip_removals, aes(x=Port_Of_Landing, y=ntrips)) +
  geom_col() +
  ggtitle("Number of Trips with MidTrip Removals") +
  theme(axis.text.x = element_text(angle=90, hjust=1))
ggplot(count_midtrip_removals, aes(x=Port_Of_Landing, y=ptrips)) +
  geom_col() +
  ggtitle("Proportion of Trips with MidTrip Removals")+
  theme(axis.text.x = element_text(angle=90, hjust=1))
```
<br>

#### Of trips with midtrip removals, how many are marked as within the Port of Landing?

Get vms records for trips with midtrip removals
```{r}
midtrip_removals <- filter(midtrip_removals, midtrip == "y")
midtrip_removals_vms <- left_join(midtrip_removals, dcrb_vms_part_rm, by=c("Rec_ID", "Port_Of_Landing"))
```
<br>

mark in-port records based on whether the in-port designation matches the port of landing
```{r}
midtrip_removals_vms$Port_Of_Landing <- as.character(midtrip_removals_vms$Port_Of_Landing)
midtrip_removals_vms <- midtrip_removals_vms %>%
  mutate(midreturn = ifelse(is.na(in_port), NA,
                            ifelse(Port_Of_Landing == in_port & to_rm == "remove", "Y", "N")))
```
<br>

count trips where midtrip removals were returns to Port of Landing
```{r}
midtrip_returns <- midtrip_removals_vms %>%
  group_by(Port_Of_Landing, Rec_ID) %>%
  summarise(didreturn = ifelse(sum(midreturn == "Y", na.rm=TRUE) > 2, "Y", "N"))
count_midtrip_returns <- midtrip_returns %>%
  group_by(Port_Of_Landing) %>%
  summarise(ntrips = sum(didreturn == "Y"), ptrips = ntrips / n())
```
<br>

```{r}
ggplot(count_midtrip_returns, aes(x=Port_Of_Landing, y=ntrips)) +
  geom_col() +
  ggtitle("Number of Trips with MidTrip Removals\n that Returned to Port of Landing") +
  theme(axis.text.x = element_text(angle=90, hjust=1))
ggplot(count_midtrip_returns, aes(x=Port_Of_Landing, y=ptrips)) +
  geom_col() +
  ggtitle("Proportion of Trips with MidTrip Removals\n that Returned to Port of Landing")+
  theme(axis.text.x = element_text(angle=90, hjust=1))
```
<br> 

#### Of all trips, what proportion are returning to the Port of Landing?

Get vms records for trips with midtrip removals
```{r}
midtrip_removals <- filter(midtrip_removals, midtrip == "y")
midtrip_removals_vms <- left_join(midtrip_removals, dcrb_vms_part_rm, by=c("Rec_ID", "Port_Of_Landing"))
```
<br>

mark in-port records based on whether the in-port designation matches the port of landing
```{r}
midtrip_removals_vms$Port_Of_Landing <- as.character(midtrip_removals_vms$Port_Of_Landing)
midtrip_removals_vms <- midtrip_removals_vms %>%
  mutate(midreturn = ifelse(is.na(in_port), NA,
                            ifelse(Port_Of_Landing == in_port, "Y", "N")))
```
<br>

count trips where midtrip removals were returns to Port of Landing
```{r}
midtrip_returns_total <-  dcrb_vms_part_rm %>%
  dplyr::select(Port_Of_Landing, Rec_ID) %>%
  distinct() %>%
  mutate(didreturn = ifelse(Rec_ID %in% unique(midtrip_returns$Rec_ID), "Y", "N"))
count_midtrip_returns_total <- midtrip_returns_total %>%
  group_by(Port_Of_Landing) %>%
  summarise(ntrips = sum(didreturn == "Y"), ptrips = ntrips / n())
```
<br>

```{r}
ggplot(count_midtrip_returns_total, aes(x=Port_Of_Landing, y=ntrips)) +
  geom_col() +
  ggtitle("Number of Trips that Returned to Port of Landing") +
  theme(axis.text.x = element_text(angle=90, hjust=1))
ggplot(count_midtrip_returns_total, aes(x=Port_Of_Landing, y=ptrips)) +
  geom_col() +
  ggtitle("Proportion of Trips that Returned to Port of Landing")+
  theme(axis.text.x = element_text(angle=90, hjust=1))
```
<br>

#### Take a look at some of these midtrip returns
```{r}
midtrip_returns_vms <-  dcrb_vms %>%
  filter(Rec_ID %in% unique(midtrip_returns$Rec_ID))
```
<br>

```{r}
plot_ports <- unique(midtrip_returns_vms$Port_Of_Landing)
for(p in plot_ports){
  tmp_vms <- filter(midtrip_returns_vms, Port_Of_Landing == p)
  if(length(unique(tmp_vms$Rec_ID)) > 2){
    trips_to_keep <- unique(tmp_vms$Rec_ID)[1:3]
    for(i in trips_to_keep){
      tmp_dat <- filter(tmp_vms, Rec_ID == i)
      out_dat <- filter(dcrb_vms_part_rm, Port_Of_Landing == p & Rec_ID == i)
      label_dat <- rbind(tmp_dat[1,], tmp_dat[length(tmp_dat$Rec_ID),])
      myplot <- ggplot(tmp_dat) +
        geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
        geom_point(aes(x=LONGITUDE, y=LATITUDE, col=to_rm), size = 4, alpha = 0.5) + 
        geom_path(aes(x=LONGITUDE, y=LATITUDE), col="dodgerblue4", size = 0.5, alpha = 0.5) +
        geom_text(data=label_dat, aes(y = LATITUDE,x=LONGITUDE, label=c("S", "E"))) +
        geom_point(data=filter(ports_df, as.character(port_code) == as.character(unique(tmp_dat$Port_Of_Landing))), aes(x=Lon, y=Lat), col="red", size=2) +
        #scale_color_continuous() +
        ggtitle(paste0("Trip ID:", i, "Port:", tmp_dat$Port_Of_Landing[1])) +
        coord_cartesian(xlim=c(min(tmp_dat$LONGITUDE) - 0.1, max(tmp_dat$LONGITUDE) + 0.1), ylim=c(min(tmp_dat$LATITUDE) - 0.1, max(tmp_dat$LATITUDE) + 0.1))
      print(myplot)
      png(paste0("ProcessVMS/scripts/quality/midtrip_returns/DCRB_VMS_", i, "_Port", tmp_dat$Port_Of_Landing[1], ".png"))
      print(myplot)
      dev.off()
      write.csv(x=out_dat, paste0("ProcessVMS/scripts/quality/midtrip_returns/DCRB_VMS_", i, "_Port", tmp_dat$Port_Of_Landing[1], ".csv"))
    }
  } else if(length(unique(tmp_vms$Rec_ID)) > 0){
    trips_to_keep <- unique(tmp_vms$Rec_ID)
    for(i in trips_to_keep){
      tmp_dat <- filter(tmp_vms, Rec_ID == i)
      out_dat <- filter(dcrb_vms_part_rm, Port_Of_Landing == p & Rec_ID == i)
      label_dat <- rbind(tmp_dat[1,], tmp_dat[length(tmp_dat$Rec_ID),])
      myplot <- ggplot(tmp_dat) +
        geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
        geom_point(aes(x=LONGITUDE, y=LATITUDE, col=to_rm), size = 4, alpha = 0.5) + 
        geom_path(aes(x=LONGITUDE, y=LATITUDE), col="dodgerblue4", size = 0.5, alpha = 0.5) +
        geom_text(data=label_dat, aes(y = LATITUDE,x=LONGITUDE, label=c("S", "E"))) +
        geom_point(data=filter(ports_df, as.character(port_code) == as.character(unique(tmp_dat$Port_Of_Landing))), aes(x=Lon, y=Lat), col="red", size=2) +
        #scale_color_continuous() +
        ggtitle(paste0("Trip ID:", i, " Port:", tmp_dat$Port_Of_Landing[1])) +
        coord_cartesian(xlim=c(min(tmp_dat$LONGITUDE) - 0.1, max(tmp_dat$LONGITUDE) + 0.1), ylim=c(min(tmp_dat$LATITUDE) - 0.1, max(tmp_dat$LATITUDE) + 0.1))
      print(myplot)
      png(paste0("ProcessVMS/scripts/quality/midtrip_returns/DCRB_VMS_", i, "_Port", tmp_dat$Port_Of_Landing[1], ".png"))
      print(myplot)
      dev.off()
      write.csv(x=out_dat, paste0("ProcessVMS/scripts/quality/midtrip_returns/DCRB_VMS_", i, "_Port", tmp_dat$Port_Of_Landing[1], ".csv"))
    }
  } else{cat("No VMS records for port: ", p, "\n")}
}
```



















