---
title: "Change in Price Per Pound"
output: html_document
---

How did price per pound change change by month?

- Time period: 2009-17, monthly

- Spatial domains: California, central vs northern CA (S of BDG, BDG and N)

- Especially with KFs observation that pots in water much longer in spring 2014 in California than in typical years.



Steps in the script: 

- Grab all DCRB fish tickets
- Get mean & sd of ppp for Dcrab for each port group, in each month. 
- Get mean & sd of ppp for Dcrab for each region, in each month



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "E:/Networks/Participation_Networks")
```


Load packages
```{r}
library(dplyr)
library(ggplot2)
library(foreign)
library(lubridate)
```
<br>


### Read in data
Processed fish ticket data, with lengths (2010-2018)
```{r}
years <- seq(2009,2017)
for(y in years){
  tmptix <- read.csv(paste0("input_data/processed/fish_tickets_",y,"_processed_for_networks.csv"))
  if(y==years[1]){
    tickets <- tmptix
  } else{
    if(y != 2018){
      tickets <- rbind(tickets,tmptix)
    } else{
      tmptix <- dplyr::select(tmptix, colnames(tickets))
      tickets <- rbind(tickets,tmptix)
    }
  }
}
head(tickets)
```

### Filtering
```{r}
dcrb_tickets <- tickets %>%
  filter(spid_recode=="DCRB") %>%
  filter(removal_type == "COMMERCIAL (NON-EFP)" | removal_type == "COMMERCIAL(DIRECT SALES)" | removal_type == "UNKNOWN") %>%
  filter(agid == "C")

dcrb_tickets_thin <- dplyr::select(dcrb_tickets, trip_id, year, tdate, pcgroup, spid_recode, ppp)
```

### Summarising
```{r}
dcrb_tickets_thin$tdate <- ymd(dcrb_tickets_thin$tdate)
dcrb_tickets_thin <- dcrb_tickets_thin %>%
  mutate(tmonth = month(tdate))
```
```{r}
dcrb_summary_ca <- dcrb_tickets_thin %>%
  group_by(year, tmonth) %>%
  summarise(mean_ppp = mean(ppp, na.rm=TRUE), sd_ppp = sd(ppp, na.rm=TRUE))
dcrb_summary_region <- dcrb_tickets_thin %>%
  mutate(region=ifelse(pcgroup %in% c("CCA", "ERA", "BGA", "BDA"), "North", "South")) %>%
  group_by(year, tmonth, region) %>%
  summarise(mean_ppp = mean(ppp, na.rm=TRUE), sd_ppp = sd(ppp, na.rm=TRUE))
```


### Plotting
```{r}
png("../../VMS-repo/CA_DCRB_ppp_annual.png")
ggplot(dcrb_summary_ca, aes(x=tmonth,y=mean_ppp)) +
  geom_point() +
  geom_errorbar(aes(ymin=mean_ppp-sd_ppp, ymax=mean_ppp+sd_ppp)) +
  geom_line() +
  facet_wrap(~year) +
  scale_x_continuous(breaks=seq(1,12,by=2), labels=seq(1,12,by=2)) +
  ggtitle("D.crab Reported Price per Pound\nCalifornia coastwide") +
  xlab("Ticket Month") +
  ylab("Mean PPP")
dev.off()
```



```{r}
png("../../VMS-repo/CA_DCRB_ppp_annual_byregion.png")
ggplot(dcrb_summary_region, aes(x=tmonth,y=mean_ppp, col=region)) +
  geom_point() +
  geom_errorbar(aes(ymin=mean_ppp-sd_ppp, ymax=mean_ppp+sd_ppp)) +
  geom_line() +
  facet_wrap(~year) +
  scale_x_continuous(breaks=seq(1,12,by=2), labels=seq(1,12,by=2)) +
  ggtitle("D.crab Reported Price per Pound\nCalifornia coastwide") +
  xlab("Ticket Month") +
  ylab("Mean PPP")
dev.off()
```
<br>

### By port group, 2014
```{r}
dcrb_summary_2014 <- dcrb_tickets_thin %>%
  filter(year==2015) %>%
  group_by(year, tmonth, pcgroup) %>%
  summarise(mean_ppp = mean(ppp, na.rm=TRUE), sd_ppp = sd(ppp, na.rm=TRUE))
```

```{r}
png("../../VMS-repo/CA_DCRB_ppp_annual_2015_byport.png")
ggplot(dcrb_summary_2014, aes(x=tmonth,y=mean_ppp)) +
  geom_point() +
  geom_errorbar(aes(ymin=mean_ppp-sd_ppp, ymax=mean_ppp+sd_ppp)) +
  geom_line() +
  facet_wrap(~pcgroup) +
  scale_x_continuous(breaks=seq(1,12,by=2), labels=seq(1,12,by=2)) +
  ggtitle("D.crab Reported Price per Pound\n2015") +
  xlab("Ticket Month") +
  ylab("Mean PPP")
dev.off()
```