---
title: "Check 24HR Gap Outliers"
output: html_document
---



```{r "setup", include=FALSE}
knitr::opts_knit$set(root.dir = "D:/VMS_RAship") 
```
<br>
<br>

### Prep for running code
Clear workspace
```{r}
rm(list=ls())
```
<br>

Install packages
```{r}
library(lubridate)
library(ggmap)
library(dplyr)
library(ggplot2)
library(reshape2)
```
```{r}
fishtixdir <- "R_Output/FishTix_VMS_Match_v2/v2.0/"
years <- seq(2009,2016)
seasons <- c("09-10", "10-11", "11-12", "12-13", "13-14", "14-15","15-16")
size_cutoff <- 40
states <- c("C")
targets <- c("DCRB")
```

<br>

## Make `ltraj` object
### Read in data

```{r}
years = seq(2009,2016)
vms_out <- read.csv(paste0("R_Output/FishTix_VMS_Match_v2/v2.0/cleaned_vms_portDist/VMS_Outputs_wTARGET_10d_lookback_",years[1],"_cleaned.csv"))

if(length(years) > 1){
  for(i in seq(2,length(years))){
    vms_out_tmp <- read.csv(paste0("R_Output/FishTix_VMS_Match_v2/v2.0/cleaned_vms_portDist/VMS_Outputs_wTARGET_10d_lookback_",years[i],"_cleaned.csv"))
    vms_out <- rbind(vms_out, vms_out_tmp)
  }
}

dim(vms_out)
vmsdat <- vms_out
```


Then the associated fish ticket data
```{r}
fishtix <- read.csv(paste0(fishtixdir,"FishTix_wVMS_10d_lookback_",years[1],".csv"))
fishtix$date <- date(parse_date_time(fishtix$date, orders=c("Ymd", "mdY")))

if(length(years) > 1){
  for(i in seq(2,length(years))){
    fishtix_tmp <- read.csv(paste0(fishtixdir,"FishTix_wVMS_10d_lookback_",years[i],".csv"))
    fishtix_tmp$date <- date(parse_date_time(fishtix_tmp$date, orders=c("Ymd", "mdY")))
    fishtix <- rbind(fishtix, fishtix_tmp)
  }
}
```
<br>



### Prep VMS data

Match vessel lengths / agency code to VMS records
```{r join_raw}
vmsdat$DOCNUM <- as.character(vmsdat$DOCNUM)
fishtix$drvid <- as.character(fishtix$drvid)
vms_w_landings <- left_join(vmsdat, fishtix, by=c("Rec_ID" = "Rec_ID", "DOCNUM" = "drvid"))
```
```{r thin_vms}
cols_to_keep <- c("Rec_ID","UTCDATETIM", "LATITUDE", "LONGITUDE", "AVG_SPEED", "AVG_COURSE", "DOCNUM", "DECLARATIO", "NGDC_M", "X_COORD", "Y_COORD", "VMS_RECNO", 
                  "westcoastdate", "date","Week", "year_Wk", "year",
                  "TARGET_max", "Port_Of_Landing", "port_group_code",
                  "VMS_lookback_pres",
                  "agency_code","FINAL_LENGTH", "HISTORIC_DATA")
vms_w_landings_thin <- select(vms_w_landings, cols_to_keep)
```
<br>

#### Filter for only those fish tickets from the state(s) and with the target species that you want to look at
```{r}
vms_w_landings_thin_filtered <- filter(vms_w_landings_thin, agency_code %in% states & TARGET_max %in% targets)
```
```{r}
unique(vms_w_landings_thin_filtered$agency_code)
unique(vms_w_landings_thin_filtered$TARGET_max)
```
<br>

Add in vessel category, season, and period columns
```{r mutate_size_season}
vmsdat2 <- vms_w_landings_thin_filtered %>%
  mutate(vessel_cat = ifelse(FINAL_LENGTH > size_cutoff, "Large", "Small")) %>%
  mutate(season = ifelse((year == 2009 & Week < 46), "08-09",
                         ifelse((year==2009 & Week > 45) | (year==2010 & Week < 46), "09-10", 
                            ifelse((year==2010 & Week > 45) | (year==2011 & Week < 46), "10-11",
                                  ifelse((year==2011 & Week > 45) | (year==2012 & Week < 46), "11-12",
                                         ifelse((year==2012 & Week > 45) | (year==2013 & Week < 46), "12-13",
                                                ifelse((year==2013 & Week > 45) | (year==2014 & Week < 46), "13-14",
                                                       ifelse((year==2014 & Week > 45) | (year==2015 & Week < 46), "14-15",
                                                              ifelse((year==2015 & Week > 45) | (year==2016 & Week < 46), "15-16",
                                                                     ifelse((year==2016 & Week > 45) | (year==2017 & Week < 46), "16-17",NA)))))))))) %>%
  mutate(period = ifelse(season == "15-16", "A1",
                         ifelse(season == "16-17", "A2", "Before")))
```
<br>

#### Filter for fish tickets filled out within the season start/end dates

function to determine whether a given fish ticket was recorded in or outside of CA DCRB fishing seasons.
```{r}
within_season <- function(mydates, port_groups){
  output_vec <- c()
  groupA <- c("CCA", "ERA", "BGA")
  groupB <- c("BDA", "SFA", "MNA", "MRA", "SBA", "SDA", "LAA")
  for(i in seq(1, length(mydates))){
    tmp_date <- mydates[i]
    tmp_year <- year(tmp_date)
    tmp_group <- port_groups[i]
    if(month(tmp_date) < 8){
      start_yr <- tmp_year - 1
      end_yr <- tmp_year}
    else{
      start_yr <- tmp_year
      end_yr <- tmp_year + 1
    }
    if(tmp_group %in% groupA){
      start_date <- ymd(paste0(start_yr, "-12-01"))
      end_date <- ymd(paste0(end_yr, "-07-15"))
      if((start_date < tmp_date) == TRUE & (tmp_date < end_date) == TRUE){
        output <- "Y"} else{ output <- "N"}
    } else if(tmp_group %in% groupB){
      start_date <- ymd(paste0(start_yr, "-11-15"))
      end_date <- ymd(paste0(end_yr, "-06-30"))
      if((start_date <= tmp_date) == TRUE & (tmp_date <= end_date) == TRUE){
        output <- "Y"} else{ output <- "N"}
    } else{
      print("port group not recognized")
      output <- NA
    }
    output_vec[i] <- output
  }
  return(output_vec)
}
```
<br>

Add a new column that says whether the VMS record is associated with a fish ticket that was within or outside of normal season dates
```{r}
vmsdat2 <- vmsdat2 %>%
  mutate(in_season = ifelse(within_season(date, port_group_code) == "Y", "Yes", "No"))
```
<br>



**Filter for desired seasons / records within season dates**
```{r}
vmsdat2_seasonFilter <- vmsdat2 %>%
  filter(season %in% seasons) %>%
  filter(in_season == "Yes") %>%
  arrange(desc(westcoastdate))
```
```{r}
dim(vmsdat2_seasonFilter)[1] / dim(vmsdat2)[1] 
```
<br>





#### Does one vessel have two size classifications in the same season? 
```{r}
vessel_categories <- vmsdat2_seasonFilter %>%
  select(DOCNUM, season, vessel_cat) %>%
  distinct() %>%
  group_by(DOCNUM, season) %>%
  summarise(num_cats = n())
```
<br>

If there are one or more vessels that have two size classifications in the same season...
```{r}
if(sum(vessel_categories$num_cats == 2) > 0){
  ## Find the data for the vessel(s) that have different size categories for the same season. What are the different lengths?
  tmp_vessels <- vessel_categories[which(vessel_categories$num_cats==2),]
  tmp_vessels_info <- left_join(tmp_vessels, vmsdat2_seasonFilter, by=c("DOCNUM", "season"))
  tmp_vessels_info %>% select(DOCNUM, season, FINAL_LENGTH) %>% distinct
  
  ## get the number of tickets per vessel category within each season
  tmp_vessels_tickets_per_cat <- tmp_vessels_info %>%
  group_by(DOCNUM, season, vessel_cat) %>%
    summarise(n_per_cat = length(unique(Rec_ID)))
  tmp_vessels_tickets_per_cat
  
  ## plot it
  myplot <- ggplot(tmp_vessels_tickets_per_cat, aes(x=DOCNUM, y=n_per_cat)) +
    geom_col(aes(fill=vessel_cat)) +
    facet_wrap(~season) +
    ylab("Number of Trips per Category")
  print(myplot)
  
  ## pull out the maximum number of tickets recorded
  max_tickets_per_cat <- tmp_vessels_tickets_per_cat %>%
    group_by(DOCNUM) %>%
    filter(n_per_cat == max(n_per_cat))
  
  ## reassign vessel category for each vessel
  for(i in seq(1:length(max_tickets_per_cat$DOCNUM))){
    tmp_vessel <- max_tickets_per_cat[i,]
    tmp_docnum <- tmp_vessel$DOCNUM
    tmp_season <- tmp_vessel$season
    cat("Reassigned vessel category (as ", tmp_vessel$vessel_cat, ") to vessel", tmp_docnum, " for season", tmp_season, "\n")
    vmsdat2_seasonFilter$vessel_cat[which(vmsdat2_seasonFilter$DOCNUM == tmp_docnum & vmsdat2_seasonFilter$season == tmp_season)] <- tmp_vessel$vessel_cat
  }
  
  
  ## did it work?
  reassigned_vessels <- vmsdat2_seasonFilter %>%
    filter(DOCNUM %in% tmp_vessels_tickets_per_cat$DOCNUM & season %in% tmp_vessels_tickets_per_cat$season) %>%
    select(DOCNUM, season, FINAL_LENGTH, vessel_cat) %>%
    distinct() %>%
    arrange(DOCNUM)
  reassigned_vessels
} else{
  print("No vessels assigned to different length categories within the same season!")
}
```
<br>



### Create data frame for `as.ltraj` function, with trip as "id"

#### Ensure that R recognises the timestamp as a POSIXct class
```{r}
vmsdat2_seasonFilter$westcoastdate <- ymd_hms(vmsdat2_seasonFilter$westcoastdate, tz= "America/Los_Angeles")
class(vmsdat2_seasonFilter$westcoastdate)
vmsdat2_seasonFilter$westcoastdate <- ymd_hms(vmsdat2_seasonFilter$westcoastdate, tz= "America/Los_Angeles")
class(vmsdat2_seasonFilter$westcoastdate)
```
<br>


#### Deal with duplicates
Remove all rows where "VMS_lookback_pres" is recorded as "Duplicated". *These Rec_IDs will be duplicates, so will return an error when creating an ltraj object*.
```{r}
dim(vmsdat2_seasonFilter)
vmsdat2_seasonFilter_nodup1 <- filter(vmsdat2_seasonFilter, VMS_lookback_pres != "Duplicated")
dim(vmsdat2_seasonFilter_nodup1)
cat("Proportion of VMS records removed due to duplicated fish tickets:", 1-dim(vmsdat2_seasonFilter_nodup1)[1]/dim(vmsdat2_seasonFilter)[1])
```
<br>

Ensure that there are no more duplicated dates under the same Rec_ID
```{r}
dim(vmsdat2_seasonFilter_nodup1)
vmsdat2_seasonFilter_nodup <- vmsdat2_seasonFilter_nodup1[!duplicated(vmsdat2_seasonFilter_nodup1[,c("Rec_ID","westcoastdate", "UTCDATETIM")]),]
dim(vmsdat2_seasonFilter_nodup)

cat("Proportion of VMS records removed for being true duplicate records:", 1-dim(vmsdat2_seasonFilter_nodup)[1]/dim(vmsdat2_seasonFilter_nodup1)[1])
```
<br>



#### Write file for use in future
Create a file with the input data frame for the `as.ltraj` function.
```{r}
write.csv(x=vmsdat2_seasonFilter_nodup, "R_Output/FishTix_VMS_Match_v2/cleaned_vms/VMSdat_for_ltraj_2009-16_CA_targetDCRB_inseason_SeasonWeek_all_perTrip.csv", row.names=FALSE, quote = FALSE)
```
<br>




### Create `ltraj` object, with trip as "id"
Load package (note-loading package in the beginning of the script will mask "id" variable from dplyr.)
```{r}
library(adehabitatLT)
library(dplyr)
```

```{r}
vmsdat2_seasonFilter_nodup$DECLARATIO[vmsdat2_seasonFilter_nodup$DECLARATIO == "N/A"] <- NA
colnames(vmsdat2_seasonFilter_nodup)
```
<br>


Create an `ltraj` object
```{r}
ltraj <- adehabitatLT::as.ltraj(xy=vmsdat2_seasonFilter_nodup[,c("X_COORD","Y_COORD")], 
                date=vmsdat2_seasonFilter_nodup$westcoastdate, 
                id=vmsdat2_seasonFilter_nodup$Rec_ID,
                infolocs = vmsdat2_seasonFilter_nodup[,c("DOCNUM",
                                        "vessel_cat", "season", "period",
                                        "westcoastdate", "VMS_RECNO", "port_group_code",
                                        "year", "date")])
```
<br>


Turn it back into a data frame
```{r}
vmsdat_ltraj_expanded <- ld(ltraj)
```

<br>

Edit the column names for the fish ticket `date`
```{r}
colnames(vmsdat_ltraj_expanded)[length(colnames((vmsdat_ltraj_expanded)))] <- "ticket_date"
colnames(vmsdat_ltraj_expanded)[which(colnames(vmsdat_ltraj_expanded) == "date.x")] <- "date"
```
<br>

Write out `ltraj` data frame.
```{r}
write.csv(vmsdat_ltraj_expanded, "scripts/final_vms/time_gaps/VMSdat_cleaned_ltraj_2009-16_CA_targetDCRB_perTrip.csv", row.names=FALSE)
```
<br>






### Identify outliers based on 24hr gaps

Find the maximum dt between VMS points
```{r}
tripdat <- vmsdat_ltraj %>%
  group_by(id, port_group_code, period) %>%
  summarise(max_dt = max(dt, na.rm=TRUE)/3600)
head(tripdat)
```


```{r}
tripdat_before <- tripdat %>%
  filter(period == "Before") %>%
  filter(!is.na(max_dt) & max_dt != 0) %>%
  select(-period)
tripdat_a1 <- tripdat %>%
  filter(period == "A1") %>%
  filter(!is.na(max_dt) & max_dt != 0) %>%
  select(-period)
```

boxplots
```{r}
tripdat_before_melt <- melt(tripdat_before, id.vars = c("id", "port_group_code"), measure.vars = c("max_dt"))
ggplot(tripdat_before_melt, aes(x=variable, y=value)) +
  geom_boxplot() +
  facet_wrap(~port_group_code) +
  ylab("Maximum time diff. between VMS (hrs)") +
  xlab("port group code") +
  ggtitle("pre-2015")
```


```{r}
tripdat_a1_melt <- melt(tripdat_a1, id.vars = c("id", "port_group_code"), measure.vars = c("max_dt"))
ggplot(tripdat_a1_melt, aes(x=variable, y=value)) +
  geom_boxplot() +
  facet_wrap(~port_group_code) +
  ylab("Maximum time diff. between VMS (hrs)")+
  xlab("port group code") +
  ggtitle("2015-16")
```





### Identify dt > 24 hours


Get the trip IDs for the outliers
```{r}
outliers_df <- tripdat_before %>%
  filter(max_dt > 72)

outliers_before <- left_join(outliers_df, vms_out, by=c("id"="Rec_ID"))
```

```{r}
outliers_df <- tripdat_a1 %>%
  filter(max_dt > 72)

outliers_a1 <- left_join(outliers_df, vms_out, by=c("id"="Rec_ID"))
```

```{r}
dim(outliers_a1 %>% select(id, port_group_code) %>% distinct()) #140
dim(tripdat_a1%>% select(id, port_group_code) %>% distinct()) #1291
dim(outliers_before %>%select(id, port_group_code) %>% distinct()) #1519
dim(tripdat_before %>% select(id, port_group_code) %>% distinct()) #12707
```

### plot the trips with outliers

```{r}
library(maps)
library(maptools)
library(ggrepel)
library(ggplot2)



USA <- map_data("world") %>% 
  filter(region=="USA")
states <- map_data("state")
west_coast <- subset(states, region %in% c("california", "oregon", "washington"))

library(RColorBrewer)
for(i in unique(outliers_a1$id)){
  tmp_dat <- dplyr::filter(outliers_a1, id==i)
  myplot <- ggplot() +
    geom_polygon(data = states, aes(x=long, y = lat, fill=region, group=group), color=" grey37", alpha=0.3) +
    guides(fill=FALSE)+
    geom_point(data=tmp_dat, aes(x=LONGITUDE, y=LATITUDE, col=westcoastdate_notime), size = 2, alpha = 0.5) + 
    scale_color_brewer(palette="Blues") +
    ggtitle(paste0("Trip ID:", i)) +
    coord_map(xlim= c(-115, -135),  ylim = c(30,45))
  print(myplot)
  png(paste0("scripts/final_vms/time_gaps/Dt72hr_A1_VMStracks_", i, ".png"))
  print(myplot)
  dev.off()
}

```






