---
title: "Plot Trip Tracks against Port of Landing"
output: html_document
---

While checking the location of the first VMS point against the previous port of landing, I found one trip where the VMS tracks were very far away from the recorded port of landing. I'm worried that this is a pervasive problem in the data set as a result of a glitch in the mapping script. 


```{r "setup", include=FALSE}
knitr::opts_knit$set(root.dir = "E:/VMS_RAship") 
```
<br>

Install packages
```{r}
library(lubridate)
library(ggmap)
library(dplyr)
library(ggplot2)
```


### Read in the data

Fish tickets matched for 2010
```{r}
fishtix <- read.csv("R_Output/FishTix_VMS_Match_v2/FishTix_wVMS_10d_lookback_2010.csv")
```
<br>

Filter out all duplicated fish tickets
```{r}
fishtix_nodup <- filter(fishtix, VMS_lookback_pres != "Duplicated")
```




VMS data matched for 2010
```{r}
vmsdat <- read.csv("R_Output/FishTix_VMS_Match_v2/cleaned_vms/VMS_Outputs_wTARGET_10d_lookback_2010_cleaned.csv")
```



### Filter for tickets from large vessels that targeted DCRB, and were landed in california
```{r}
fishtix_filtered <- fishtix_nodup %>%
  filter(agency_code == "C") %>%
  filter(TARGET=="DCRB") %>%
  filter(FINAL_LENGTH >= 40) %>%
  filter(VMS_lookback_pres==1)
```





### Add columns for lat/long of port of landing

```{r}
ports_landed <- data.frame(ports=unique(fishtix_filtered$pacfin_port_code))
towns <- c("San Francisco", "Half Moon Bay", "Bodega Bay", "Santa Cruz", "Crescent City", 
           "Eureka", "San Francisco Area",
           "Fort Bragg", "Point Arena",
           "Sausalito")
port_lat <- c(37.774929, 37.463638, 38.317451, 36.974117, 41.852807,
         40.786140, 37.774929,
         39.453473, 38.90880,
         37.859027)
port_long <- c(-122.419418, -122.429237, -123.028648, -122.030792, -124.149862,
          -124.161308, -122.419418,
          -123.814911, -123.69307,
          -122.485469)

port_x <- c(551129, 550532, 495800, 586266, 400024,
            401840, 551129,
            430708, 439909,
            545280)
port_y <- c(4181002, 4146451, 4242792, 4092440, 4623357,
            4517437, 4181002,
            4366550, 4306884,
            4190306)


port_landed_coords <- cbind(ports_landed, towns, port_lat, port_long, port_x, port_y)
```



### Join to VMS data 

```{r}
fishtix_allVMS <- left_join(fishtix_filtered, vmsdat, by=c("drvid" = "DOCNUM", "Rec_ID"="Rec_ID"))
```
```{r}
fishtix_allVMS <- left_join(fishtix_allVMS, port_landed_coords, by=c("pacfin_port_code" = "ports"))
```


### Plot

```{r}
library(tidyverse)
library(maps)
library(maptools)
library(ggrepel)
library(viridis)
library(ggplot2)
library(pals)
library(grid)
```


```{r}
USA <- map_data("world") %>% 
  filter(region=="USA")
states <- map_data("state")
west_coast <- subset(states, region %in% c("california", "oregon", "washington"))
```

```{r}
for(id in unique(fishtix_allVMS$Rec_ID)){
  tmp_dat <- dplyr::filter(fishtix_allVMS, Rec_ID==id)
  myplot <- ggplot() +
    geom_polygon(data = states, aes(x=long, y = lat, fill=region, group=group), color=" grey37", alpha=0.3) +
    guides(fill=FALSE)+
    geom_point(data=filter(port_landed_coords, ports==unique(tmp_dat$pacfin_port_code)), aes(x=port_long, y=port_lat), size=2, col="red") +
    geom_point(data=tmp_dat, aes(x=LONGITUDE, y=LATITUDE), size = 2, alpha = 0.7, col="blue") + 
    coord_map(xlim= c(-120, -125),  ylim = c(30,45)) +
    ggtitle(paste0("Trip ID:", id))
  print(myplot)
}
```


```{r}
id=48701427
tmp_dat <- dplyr::filter(fishtix_allVMS, Rec_ID==id)
myplot <- ggplot() +
  geom_polygon(data = states, aes(x=long, y = lat, fill=region, group=group), color=" grey37", alpha=0.3) +
  guides(fill=FALSE)+
  geom_point(data=filter(port_landed_coords, ports==unique(tmp_dat$pacfin_port_code)), aes(x=port_long, y=port_lat), size=2, col="red") +
  geom_point(data=tmp_dat, aes(x=LONGITUDE, y=LATITUDE), size = 2, alpha = 0.7, col="blue") + 
  coord_map(xlim= c(-120, -125),  ylim = c(30,45)) +
  ggtitle(paste0("Trip ID:", id))
print(myplot)
```






