---
title: "Check Max Net Displacement Outliers"
output: html_document
---


```{r "setup", include=FALSE}
knitr::opts_knit$set(root.dir = "D:/VMS_RAship") 
```
<br>
<br>

### Prep for running code
Clear workspace
```{r}
rm(list=ls())
```
<br>

Install packages
```{r}
library(lubridate)
library(ggmap)
library(dplyr)
library(ggplot2)
library(reshape2)
```


<br>

### Read in data

`Ltraj` data frame
```{r}
vmsdat_ltraj <- read.csv("results/Participation_by_PortGroup/VMSdat_ltraj_2009-16_CA_targetDCRB_inseason_SeasonWeek_all_perTrip.csv")
```
```{r}
colnames(vmsdat_ltraj)
```
```{r}
colnames(vmsdat_ltraj)[which(colnames(vmsdat_ltraj)=="date.1")] <- "ticket.date"
```
<br>


Then, the full vms data set
```{r}
years = seq(2009,2016)
vms_out <- read.csv(paste0("R_Output/FishTix_VMS_Match_v2/cleaned_vms/VMS_Outputs_wTARGET_10d_lookback_",years[1],"_cleaned.csv"))

if(length(years) > 1){
  for(i in seq(2,length(years))){
    vms_out_tmp <- read.csv(paste0("R_Output/FishTix_VMS_Match_v2/cleaned_vms/VMS_Outputs_wTARGET_10d_lookback_",years[i],"_cleaned.csv"))
    vms_out <- rbind(vms_out, vms_out_tmp)
  }
}
```
```{r}
dim(vms_out)
```





### Identify outliers based on IQR, period + port group



Tukey's method removes any observation outside of the IQR in a boxplot. 
For this data, I want to remove outliers in trip distance and duration, for fish tickets landed at each port group and in each period


First, summarise the trip data
```{r}
tripdat <- vmsdat_ltraj %>%
  group_by(id, port_group_code, period) %>%
  summarise(max_disp = sqrt(max(R2n, na.rm=TRUE)))
head(tripdat)
```

```{r}
tripdat_before <- tripdat %>%
  filter(period == "Before") %>%
  filter(!is.na(max_disp) | max_disp != 0) %>%
  select(-period)
tripdat_a1 <- tripdat %>%
  filter(period == "A1") %>%
  filter(!is.na(max_disp) | max_disp != 0) %>%
  select(-period)
```

boxplots
```{r}
tripdat_before_melt <- melt(tripdat_before, id.vars = c("id", "port_group_code"), measure.vars = c("max_disp"))
ggplot(tripdat_before_melt, aes(x=variable, y=value)) +
  geom_boxplot() +
  facet_wrap(~port_group_code)
```






Get the trip IDs for the outliers
```{r}
mydat <- tripdat_before

outliers_df <- data.frame(id = as.character(),
                          max_disp = as.numeric())
for(p in unique(mydat$port_group_code)){
  tmpdat <- filter(mydat, port_group_code == p)
  outliers <- boxplot.stats(tmpdat$max_disp, do.out=TRUE)$out  # outlier values.
  ## get indices for outliers
  ids <- tmpdat$id[which(tmpdat$max_disp %in% outliers)]
  ## append to data frames
  tmp_df <- data.frame(id=ids,
                       max_disp= outliers)
  outliers_df <- rbind(outliers_df, tmp_df)
}

outliers_before <- left_join(outliers_df, vms_out, by=c("id"="Rec_ID"))
```

```{r}
mydat <- tripdat_a1

outliers_df <- data.frame(id = as.character(),
                          max_disp = as.numeric())
for(p in unique(mydat$port_group_code)){
  tmpdat <- filter(mydat, port_group_code == p)
  outliers <- boxplot.stats(tmpdat$max_disp, do.out=TRUE)$out  # outlier values.
  ## get indices for outliers
  ids <- tmpdat$id[which(tmpdat$max_disp %in% outliers)]
  ## append to data frames
  tmp_df <- data.frame(id=ids,
                       max_disp= outliers)
  outliers_df <- rbind(outliers_df, tmp_df)
}

outliers_a1 <- left_join(outliers_df, vms_out, by=c("id"="Rec_ID"))
```


### plot the trips with outliers

```{r}
library(maps)
library(maptools)
library(ggrepel)
library(ggplot2)



USA <- map_data("world") %>% 
  filter(region=="USA")
states <- map_data("state")
west_coast <- subset(states, region %in% c("california", "oregon", "washington"))

library(RColorBrewer)
for(i in unique(outliers_a1$id)){
  tmp_dat <- dplyr::filter(outliers_a1, id==i)
  myplot <- ggplot() +
    geom_polygon(data = states, aes(x=long, y = lat, fill=region, group=group), color=" grey37", alpha=0.3) +
    guides(fill=FALSE)+
    geom_point(data=tmp_dat, aes(x=LONGITUDE, y=LATITUDE, col=westcoastdate_notime), size = 2, alpha = 0.5) + 
    scale_color_brewer(palette="Blues") +
    ggtitle(paste0("Trip ID:", i)) +
    coord_map(xlim= c(-115, -135),  ylim = c(30,45))
  print(myplot)
  png(paste0("Verify_Code/Filter_DistDur_Outliers/by_port_group/MaxDisp_Outliers_VMStracks_", i, ".png"))
  print(myplot)
  dev.off()
}

```



```{r}
library(maps)
library(maptools)
library(ggrepel)
library(ggplot2)



USA <- map_data("world") %>% 
  filter(region=="USA")
states <- map_data("state")
west_coast <- subset(states, region %in% c("california", "oregon", "washington"))

library(RColorBrewer)
for(i in unique(outliers_before$id)){
  tmp_dat <- dplyr::filter(outliers_before, id==i)
  myplot <- ggplot() +
    geom_polygon(data = states, aes(x=long, y = lat, fill=region, group=group), color=" grey37", alpha=0.3) +
    guides(fill=FALSE)+
    geom_point(data=tmp_dat, aes(x=LONGITUDE, y=LATITUDE, col=westcoastdate_notime), size = 2, alpha = 0.5) + 
    scale_color_brewer(palette="Blues") +
    ggtitle(paste0("Trip ID:", i)) +
    coord_map(xlim= c(-115, -135),  ylim = c(30,45))
  print(myplot)
  png(paste0("Verify_Code/Filter_DistDur_Outliers/by_port_group/MaxDisp_Outliers_Before_VMStracks_", i, ".png"))
  print(myplot)
  dev.off()
}

```






### Identify outliers based on IQR, period




Get the trip IDs for the outliers
```{r}
mydat <- tripdat_before

outliers_df <- data.frame(id = as.character(),
                          max_disp = as.numeric())
outliers <- boxplot.stats(mydat$max_disp, do.out=TRUE)$out  # outlier values.
## get indices for outliers
ids <- mydat$id[which(mydat$max_disp %in% outliers)]
## append to data frames
tmp_df <- data.frame(id=ids,
                     max_disp= outliers)
outliers_df <- rbind(outliers_df, tmp_df)

outliers_before <- left_join(outliers_df, vms_out, by=c("id"="Rec_ID"))
```

```{r}
mydat <- tripdat_a1

outliers_df <- data.frame(id = as.character(),
                          max_disp = as.numeric())
outliers <- boxplot.stats(mydat$max_disp, do.out=TRUE)$out  # outlier values.
## get indices for outliers
ids <- mydat$id[which(mydat$max_disp %in% outliers)]
## append to data frames
tmp_df <- data.frame(id=ids,
                     max_disp= outliers)
outliers_df <- rbind(outliers_df, tmp_df)

outliers_a1 <- left_join(outliers_df, vms_out, by=c("id"="Rec_ID"))
```

```{r}
dim(outliers_a1 %>% select(id) %>% distinct()) #57
dim(tripdat_a1%>% select(id) %>% distinct()) #1260
dim(outliers_before %>%select(id) %>% distinct()) #627
dim(tripdat_before %>% select(id) %>% distinct()) #12323
```


### plot the trips with outliers

```{r}
library(maps)
library(maptools)
library(ggrepel)
library(ggplot2)



USA <- map_data("world") %>% 
  filter(region=="USA")
states <- map_data("state")
west_coast <- subset(states, region %in% c("california", "oregon", "washington"))

library(RColorBrewer)
for(i in unique(outliers_a1$id)){
  tmp_dat <- dplyr::filter(outliers_a1, id==i)
  myplot <- ggplot() +
    geom_polygon(data = states, aes(x=long, y = lat, fill=region, group=group), color=" grey37", alpha=0.3) +
    guides(fill=FALSE)+
    geom_point(data=tmp_dat, aes(x=LONGITUDE, y=LATITUDE, col=westcoastdate_notime), size = 2, alpha = 0.5) + 
    scale_color_brewer(palette="Blues") +
    ggtitle(paste0("Trip ID:", i)) +
    coord_map(xlim= c(-115, -135),  ylim = c(30,45))
  print(myplot)
  png(paste0("Verify_Code/Filter_DistDur_Outliers/MaxDisp_Outliers_A1_VMStracks_", i, ".png"))
  print(myplot)
  dev.off()
}

```



```{r}
```






