---
title: "Compare Filtered v. UnFiltered VMS"
output: html_document
---


In an effort to understand which VMS records were added into the data set by using the unfiltered VMS data, I'm comparing the new unfiltered data with the filtered data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "D:/VMS-repo") 
```
Clear workspace
```{r}
rm(list=ls())
```

```{r}
library(lubridate)
library(dplyr)
library(ggplot2)
```
<br>



### Read in
unfiltered data
```{r}
mydat <- read.csv("ProcessVMS/R_Output/match/unfiltered/cleaned/VMS_Outputs_wTARGET_10d_lookback_2010_cleaned.csv")
```

ArcGIS filtered data
```{r}
mydat_gisfilter <- read.csv("../VMS_RAship/R_Output/FishTix_VMS_Match_v2/cleaned/VMS_Outputs_wTARGET_10d_lookback_2010_cleaned.csv")
```

R-filtered
```{r}
dcrb_rfilter <- read.csv("ProcessVMS/R_Output/match/unfiltered/cleaned/DCRB_VMS_Outputs_wTARGET_10d_lookback_2010_tripedge_inport_rm.csv")
```
<br>

ltraj
```{r}
vms_ltraj <- read.csv("ProcessVMS/R_Output/interpolation/ltraj_VMS2010_LAT_CA_targetDCRB_inseason_SeasonWeek_all_perTrip.csv")
```


ArcGIS filtered ltraj
```{r}
getwd()
ltraj_gisfilter <- read.csv("ProcessVMS/R_Output/interpolation/ltraj_VMS2010_LAT_filtered_CA_targetDCRB_inseason_SeasonWeek_all_perTrip.csv")
```

Fish ticket data 
```{r}
fishtix <- read.csv("../VMS_RAship/R_Output/FishTix_VMS_Match_v2/unfiltered/FishTix_wVMS_10d_lookback_2010.csv")
```
<br>


## Prepare the `ltraj` object for the new R filtered data

Match vessel lengths / agency code to VMS records
```{r join_raw}
dcrb_rfilter$DOCNUM <- as.character(dcrb_rfilter$DOCNUM)
fishtix$drvid <- as.character(fishtix$drvid)
vms_w_landings <- left_join(dcrb_rfilter, fishtix, by=c("Rec_ID" = "Rec_ID", "DOCNUM" = "drvid"))
```
```{r thin_vms}
cols_to_keep <- c("Rec_ID","UTCDATETIM", "LATITUDE", "LONGITUDE", "AVG_SPEED", "AVG_COURSE", "DOCNUM", "DECLARATIO", "NGDC_M", "X_COORD", "Y_COORD", "VMS_RECNO", 
                  "westcoastdate", "date","Week", "year_Wk", "year",
                  "TARGET_max", "Port_Of_Landing", "port_group_code",
                  "VMS_lookback_pres",
                  "agency_code","FINAL_LENGTH", "HISTORIC_DATA")
vms_w_landings_thin <- select(vms_w_landings, cols_to_keep)
```
<br>


Add in vessel category, season, and period columns
```{r mutate_size_season}
vmsdat2 <- vms_w_landings_thin %>%
  mutate(vessel_cat = ifelse(FINAL_LENGTH > 40, "Large", "Small")) %>%
  mutate(season = ifelse((year == 2009 & Week < 46), "08-09",
                         ifelse((year==2009 & Week > 45) | (year==2010 & Week < 46), "09-10", 
                            ifelse((year==2010 & Week > 45) | (year==2011 & Week < 46), "10-11",
                                  ifelse((year==2011 & Week > 45) | (year==2012 & Week < 46), "11-12",
                                         ifelse((year==2012 & Week > 45) | (year==2013 & Week < 46), "12-13",
                                                ifelse((year==2013 & Week > 45) | (year==2014 & Week < 46), "13-14",
                                                       ifelse((year==2014 & Week > 45) | (year==2015 & Week < 46), "14-15",
                                                              ifelse((year==2015 & Week > 45) | (year==2016 & Week < 46), "15-16",
                                                                     ifelse((year==2016 & Week > 45) | (year==2017 & Week < 46), "16-17",NA)))))))))) %>%
  mutate(period = ifelse(season == "15-16", "A1",
                         ifelse(season == "16-17", "A2", "Before")))
```
<br>


Ensure that R recognises the timestamp as a POSIXct class
```{r}
vmsdat2$westcoastdate <- ymd_hms(vmsdat2$westcoastdate, tz= "America/Los_Angeles")
class(vmsdat2$westcoastdate)
```
<br>


Remove all rows where "VMS_lookback_pres" is recorded as "Duplicated". *These Rec_IDs will be duplicates, so will return an error when creating an ltraj object*.
```{r}
dim(vmsdat2)
vmsdat2_nodup1 <- filter(vmsdat2, VMS_lookback_pres != "Duplicated")
dim(vmsdat2_nodup1)
cat("Proportion of VMS records removed due to duplicated fish tickets:", 1-dim(vmsdat2_nodup1)[1]/dim(vmsdat2)[1])
```
<br>

Ensure that there are no more duplicated dates under the same Rec_ID
```{r}
dim(vmsdat2_nodup1)
vmsdat2_nodup <- vmsdat2_nodup1[!duplicated(vmsdat2_nodup1[,c("Rec_ID","westcoastdate", "UTCDATETIM")]),]
dim(vmsdat2_nodup)

cat("Proportion of VMS records removed for being true duplicate records:", 1-dim(vmsdat2_nodup)[1]/dim(vmsdat2_nodup1)[1])
```
<br>
Load package (note-loading package in the beginning of the script will mask "id" variable from dplyr.)
```{r}
library(adehabitatLT)
library(dplyr)
```

```{r}
vmsdat2_nodup$DECLARATIO[vmsdat2_nodup$DECLARATIO == "N/A"] <- NA
```



```{r}
ltraj_dcrb_rfilter <- adehabitatLT::as.ltraj(xy=vmsdat2_nodup[,c("X_COORD","Y_COORD")], 
                date=vmsdat2_nodup$westcoastdate, 
                id=vmsdat2_nodup$Rec_ID,
                infolocs = vmsdat2_nodup[,c("DOCNUM","vessel_cat",
                                                         "port_group_code", "date",
                                                         "season", "period")])
```

Turn back into a data frame
```{r}
ltraj_dcrb_rfilter <- ld(ltraj_dcrb_rfilter)
```


## Compare number of VMS records per trip

#### Over All Records

First, find number of vms records per trip
```{r}
records_unfiltered <- mydat %>%
  group_by(Rec_ID) %>%
  summarise(n_records = length(unique(VMS_RECNO))) %>%
  mutate(type="unfiltered")
records_gisfiltered <- mydat_gisfilter %>%
  group_by(Rec_ID) %>%
  summarise(n_records = length(unique(VMS_RECNO))) %>%
  mutate(type="GIS filtered")

records <- rbind(records_unfiltered, records_gisfiltered)
```

Compare histograms
```{r}
ggplot(records, aes(x=n_records, fill=type)) +
  geom_histogram() +
  facet_wrap(~type) +
  ggtitle("All Trips") +
  xlab("VMS Records per Trip") +
  ylab("Number of Trips")
```



#### Dungeness crab trips

First, find number of vms records per trip
```{r}
records_unfiltered <- mydat %>%
  filter(TARGET_max=="DCRB") %>%
  group_by(Rec_ID) %>%
  summarise(n_records = length(unique(VMS_RECNO))) %>%
  mutate(type="unfiltered")
records_gisfiltered <- mydat_gisfilter %>%
  filter(TARGET_max=="DCRB") %>%
  group_by(Rec_ID) %>%
  summarise(n_records = length(unique(VMS_RECNO))) %>%
  mutate(type="GIS filtered")
records_rfiltered <- dcrb_rfilter %>%
  group_by(Rec_ID) %>%
  summarise(n_records = length(unique(VMS_RECNO))) %>%
  mutate(type="R filtered")

records <- rbind(records_unfiltered, records_gisfiltered, records_rfiltered)
```

Compare histograms
```{r}
ggplot(records, aes(x=n_records, fill=type)) +
  geom_histogram() +
  facet_wrap(~type) +
  ggtitle("DCRB Trips") +
  xlab("VMS Records per Trip") +
  ylab("Number of Trips")
```


## Compare time gaps in single trip

#### First, all trip gaps
ltraj_gisfilter
dcrb_ltraj_rfilter
Compare histograms
```{r}
upper_lim <- c(2000, 4000, 2000, 6000, 1000, 2500, 50)
i = 1
for(p in unique(vms_ltraj$port_group_code)){
  ltraj_p <- vms_ltraj %>%
    filter(port_group_code == p) %>%
    filter(!is.na(dt)) %>%
    dplyr::select(id, dt) %>%
    mutate(type="unfiltered")
  ltraj_gisfilter_p <-  ltraj_gisfilter %>%
    filter(port_group_code == p) %>%
    filter(!is.na(dt)) %>%
    dplyr::select(id, dt) %>%
    mutate(type=" GIS filtered")
  ltraj_rfilter_p <- ltraj_dcrb_rfilter %>%
    filter(port_group_code == p) %>%
    filter(!is.na(dt)) %>%
    dplyr::select(id, dt) %>%
    mutate(type=" R filtered")
  ltraj_p <- rbind(ltraj_p, ltraj_gisfilter_p, ltraj_rfilter_p)
  myplot <- ggplot(ltraj_p, aes(x=dt/60)) +
    geom_histogram(binwidth=12) + 
    facet_wrap(~type) +
    ggtitle(paste0("Port Group: ", p)) +
    xlab("Time Between VMS Records (min)") +
    ylim(c(0,upper_lim[i]))
  print(myplot)
  i = i + 1
}
```

Compare boxplots to see outliers
```{r}
for(p in unique(vms_ltraj$port_group_code)){
  ltraj_p <- vms_ltraj %>%
    filter(port_group_code == p) %>%
    filter(!is.na(dt)) %>%
    dplyr::select(id, dt) %>%
    mutate(type="unfiltered")
  ltraj_gisfilter_p <-  ltraj_gisfilter %>%
    filter(port_group_code == p) %>%
    filter(!is.na(dt)) %>%
    dplyr::select(id, dt) %>%
    mutate(type=" GIS filtered")
  ltraj_rfilter_p <- ltraj_dcrb_rfilter %>%
    filter(port_group_code == p) %>%
    filter(!is.na(dt)) %>%
    dplyr::select(id, dt) %>%
    mutate(type=" R filtered")
  ltraj_p <- rbind(ltraj_p, ltraj_gisfilter_p, ltraj_rfilter_p)
  myplot <- ggplot(ltraj_p, aes(x=type, y=dt/3600)) +
    geom_boxplot() + 
    ggtitle(paste0("Port Group: ", p)) +
    ylab("Time between VMS Records (hours)")
  print(myplot)
}
```

  ltraj_p <- vms_ltraj %>%
    filter(port_group_code == p) %>%
    filter(!is.na(dt)) %>%
    dplyr::select(id, dt) %>%
    mutate(type="unfiltered")
  ltraj_gisfilter_p <-  ltraj_gisfilter %>%
    filter(port_group_code == p) %>%
    filter(!is.na(dt)) %>%
    dplyr::select(id, dt) %>%
    mutate(type=" GIS filtered")
  ltraj_rfilter_p <- ltraj_dcrb_rfilter %>%
    filter(port_group_code == p) %>%
    filter(!is.na(dt)) %>%
    dplyr::select(id, dt) %>%
    mutate(type=" R filtered")
  ltraj_p <- rbind(ltraj_p, ltraj_gisfilter_p, ltraj_rfilter_p)

#### Just Max trip gaps

Find the maximum dt between VMS points
```{r}
vms_ltraj$id <- as.character(vms_ltraj$id)
tripdat <- vms_ltraj %>%
  group_by(id, port_group_code, period) %>%
  summarise(max_dt = max(dt, na.rm=TRUE)/3600) %>%
  mutate(type = "unfiltered")
ltraj_gisfilter$id <- as.character(ltraj_gisfilter$id)
tripdat_gisfilter <- ltraj_gisfilter %>%
  group_by(id, port_group_code, period) %>%
  summarise(max_dt = max(dt, na.rm=TRUE)/3600) %>%
  mutate(type = "GIS filtered")
ltraj_dcrb_rfilter$id <- as.character(ltraj_dcrb_rfilter$id)
tripdat_rfilter <- ltraj_dcrb_rfilter %>%
  group_by(id, port_group_code, period) %>%
  summarise(max_dt = max(dt, na.rm=TRUE)/3600) %>%
  mutate(type = "R filtered")

tripdat <- rbind(tripdat, tripdat_gisfilter, tripdat_rfilter)
head(tripdat)
```

Compare histograms
```{r}
for(p in unique(tripdat$port_group_code)){
  tripdat_p <- filter(tripdat, port_group_code == p)
  myplot <- ggplot(tripdat_p, aes(x=max_dt)) +
    geom_histogram() + 
    facet_wrap(~type) +
    ggtitle(paste0("Port Group: ", p)) +
    xlab("Max. dt")
  print(myplot)
}
```
zooming in to the 0-24 hr range:
```{r}
for(p in unique(tripdat$port_group_code)){
  tripdat_p <- tripdat %>%
    filter(port_group_code == p) %>%
    filter(max_dt < 25)
  myplot <- ggplot(tripdat_p, aes(x=max_dt)) +
    geom_histogram() + 
    facet_wrap(~type) +
    ggtitle(paste0("Port Group: ", p)) +
    xlab("Max. dt")
  print(myplot)
}
```
<br>

