---
title: "COS Buffer Size"
output: html_document
---

Some of the COS fishing trips appear to be getting removed even though the vessel is displaced from the port location. 

<br>
```{r "setup", include=FALSE}
knitr::opts_knit$set(root.dir = "D:/VMS-repo") 
```
<br>
<br>

### Prep for running code
Clear workspace
```{r}
rm(list=ls())
```
<br>

Install packages
```{r include=FALSE}
#install.packages("foreign")
#install.packages("dplyr")
#install.packages("tidyr")
#install.packages("lubridate") #https://rstudio-pubs-static.s3.amazonaws.com/28038_1bcb9aa80ca84f27ace07d612872861a.html

library(foreign)
library(dplyr)
library(tidyr)
library(lubridate) 
library(ggplot2)
library(maps)
library(rgdal)
library(rgeos)
```
<br>

Create Objects / Data Frames
```{r}
## directory where Blake's processed VMS data is stored as .dbf files
origVMS_dir = "ProcessVMS/R_Output/match/unfiltered/v4/cleaned/"

## file with port coordinates, including directory
coords_file = "Input_Data/port_coords_fromBlake_edited.csv"

## calendar years
years <- c(2010)

## width of buffer circle (in meters). applies to all ports but those that require smaller buffers
r = 3000

## port codes for any ports that require reduced buffer sizes. Default is only COS
ports_lowbuffer <- c("COS")

## width of buffer circle (in meters) for COS
r2 = 1500
```
<br>

### Read in the data

First, the full vms data set
```{r}
vms <- read.csv(paste0(origVMS_dir,"VMS_Outputs_wTARGET_10d_lookback_2010_cleaned.csv"))
dim(vms)
head(vms)
dcrb_vms <- filter(vms, TARGET_max == "DCRB")
```



Then, the coordinates for the ports
```{r}
portdat <- read.csv(coords_file)
if(any(is.na(portdat$lat))==TRUE){
  message("WARNING: some ports are missing coordinates. ports were removed from data frame.")
  portdat <- filter(portdat, !is.na(lat))
}
port_coords <- portdat
port_coords_reg <- portdat %>% filter(!(port_code %in% ports_lowbuffer))
port_coords_sm <- portdat %>% filter(port_code %in% ports_lowbuffer)
```


### Prep the port coordinates
*Note - to assign a project, `proj4string` requires coordinates in xy (lng/lat) order*
```{r}
coordinates(port_coords_reg) <- c("Lon", "Lat") 
proj4string(port_coords_reg) <- CRS("+init=epsg:4326") # WGS 84
port_coords_reg_geo <- spTransform(port_coords_reg, CRSobj = CRS("+init=epsg:32610"))

coordinates(port_coords_sm) <- c("Lon", "Lat") 
proj4string(port_coords_sm) <- CRS("+init=epsg:4326") # WGS 84
port_coords_sm_geo <- spTransform(port_coords_sm, CRSobj = CRS("+init=epsg:32610"))
```
<br>

### Create buffer
```{r}
buffer_polygons_reg <- gBuffer(port_coords_reg_geo, byid=TRUE, width=r)
buffer_polygons_sm <- gBuffer(port_coords_sm_geo, byid=TRUE, width=r2)
buffer_polygons  <- rbind(buffer_polygons_reg, buffer_polygons_sm)
```


Get correct input format / coordinate system for ggplot
```{r}
ports_df <- data.frame(port_coords)
buffer_polygons_LL <- spTransform(buffer_polygons, CRSobj = CRS("+init=epsg:4326"))
plot(buffer_polygons_LL)
```

Basemap
```{r}
data(stateMapEnv)
states_df <- map_data("state") %>%
  filter(region == "california" | region=="oregon" | region=="washington")
```


Zoom in to a bay area and a coast area (*this is a crappy background map, so if ports are on land, that's ok.*)
```{r}
cutcoast <- list(c(-124, -124.7, 43.5, 43)) # OR [COS,GLD]
```
```{r}
for(i in c(1)){
  myplot <- ggplot() +
    geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
    geom_polygon(data=buffer_polygons_LL, aes(x = long, y = lat, group = group), col="blue") +
    geom_point(data=ports_df, aes(x=Lon, y=Lat), col="red", size=2) +
    coord_cartesian(xlim=c(cutcoast[[i]][1], cutcoast[[i]][2]), ylim=c(cutcoast[[i]][3], cutcoast[[i]][4])) +
    geom_text(data=ports_df, aes(x=Lon, y=Lat, label=port_code),size=4, nudge_x=-0.07)
  print(myplot)
}
```
<br>

### Focus in on trips from ports with smaller buffer zones

Focus in on trips leaving from COS
```{r}
dcrb_vms_sub <- filter(dcrb_vms, Port_Of_Landing %in% ports_lowbuffer)
```




### Re-filter VMS points
Assign projection to VMS points
```{r}
dcrb_vms_sp <- dcrb_vms_sub
coordinates(dcrb_vms_sp) <- c("LONGITUDE", "LATITUDE") 
proj4string(dcrb_vms_sp) <- CRS("+init=epsg:4326") # WGS 84
dcrb_vms_geo <- spTransform(dcrb_vms_sp, CRSobj = CRS("+init=epsg:32610"))
```
<br>


Identify data points within buffer zones
```{r}
in.port <- over(dcrb_vms_geo, buffer_polygons)$port_code
dcrb_vms_sub <- mutate(dcrb_vms_sub, in_port = in.port)
```
<br>

Mark "remove" if the record is in port *and* the avg speed is < 0.05.
```{r}
dcrb_vms_sub <- mutate(dcrb_vms_sub, to_rm = ifelse(!is.na(in_port) & AVG_SPEED < 1, "remove", "retain"))
cat("Proportion of VMS records removed: ", sum(dcrb_vms_sub$to_rm == "remove") / dim(dcrb_vms_sub)[1], "\n")
cat("There were ", dim(dcrb_vms_sub)[1]/sum(dcrb_vms_sub$to_rm == "retain"), "x VMS records in the unfiltered data than will be retained.\n")
cat("Proportion of VMS records marked in-port removed: ", sum(dcrb_vms_sub$to_rm == "remove") / sum(!is.na(dcrb_vms_sub$in_port)))
```
<br>

Zoom into bay with VMS points for COS trips.
```{r}
for(i in c(1)){
  myplot <- ggplot() +
    geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
    geom_polygon(data=buffer_polygons_LL, aes(x = long, y = lat, group = group), col="blue") +
    geom_point(data=ports_df, aes(x=Lon, y=Lat), col="red", size=2) +
    geom_point(data=dcrb_vms_sub, aes(x=LONGITUDE, y=LATITUDE, col=to_rm), alpha=0.5) +
    geom_text(data=ports_df, aes(x=Lon, y=Lat, label=port_code),size=4, nudge_x=0.07) +
    coord_cartesian(xlim=c(cutcoast[[i]][1], cutcoast[[i]][2]), ylim=c(cutcoast[[i]][3], cutcoast[[i]][4]))
  print(myplot)
}
```
<br>

What proportion of VMS records are being removed from COS trips, now that there is a reduced buffer zone size?
```{r}
premoved <- dcrb_vms_sub %>%
  group_by(Rec_ID, Port_Of_Landing) %>%
  summarise(p.remove = sum(to_rm=="remove") / n())
cat("Removed more than 99% of records from ", sum(premoved$p.remove > 0.99) / length(premoved$p.remove) * 100, "% (" , sum(premoved$p.remove > 0.99) , "total) of trips. ")
for(p in ports_lowbuffer){
  tmp_dat <- filter(premoved, Port_Of_Landing == p)
  myplot <- ggplot(premoved, aes(x=p.remove)) +
    geom_histogram() +
    ggtitle("Across All Ports") 
  ggplot(data=premoved, aes(x=p.remove)) +
    geom_histogram() +
    xlab("Percent VMS records removed per trip") +
    ggtitle(p)
  print(myplot)
}
```
<br>

Look at full trip data for some trips which previously had 100% records removed.
```{r}
trips_to_keep <- c(449332172, 449331741, 449332171)
for(id in trips_to_keep){
  tmp_dat <- filter(dcrb_vms_sub, Rec_ID == id)
  myplot <- ggplot(tmp_dat) +
    geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
    geom_point(aes(x=LONGITUDE, y=LATITUDE, col=to_rm), size = 4, alpha = 0.5) + 
    geom_path(aes(x=LONGITUDE, y=LATITUDE), col="dodgerblue4", size = 0.5, alpha = 0.5) +
    geom_point(data=filter(ports_df, as.character(port_code) == as.character(unique(tmp_dat$Port_Of_Landing))), aes(x=Lon, y=Lat), col="red", size=2) +
    #scale_color_continuous() +
    ggtitle(paste0("Trip ID:", i, "Port:", tmp_dat$Port_Of_Landing[1])) +
    coord_cartesian(xlim=c(min(tmp_dat$LONGITUDE) - 0.1, max(tmp_dat$LONGITUDE) + 0.1), ylim=c(min(tmp_dat$LATITUDE) - 0.1, max(tmp_dat$LATITUDE) + 0.1))
  print(myplot)
}
```
<br>


Look at full trip data for a subset of 100 trips, COS port of landing. 
```{r}
trips_to_keep <- unique(dcrb_vms_sub$Rec_ID)[1:100]
for(id in trips_to_keep){
  tmp_dat <- filter(dcrb_vms_sub, Rec_ID == id)
  myplot <- ggplot(tmp_dat) +
    geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
    geom_point(aes(x=LONGITUDE, y=LATITUDE, col=to_rm), size = 4, alpha = 0.5) + 
    geom_path(aes(x=LONGITUDE, y=LATITUDE), col="dodgerblue4", size = 0.5, alpha = 0.5) +
    geom_point(data=filter(ports_df, as.character(port_code) == as.character(unique(tmp_dat$Port_Of_Landing))), aes(x=Lon, y=Lat), col="red", size=2) +
    #scale_color_continuous() +
    ggtitle(paste0("Trip ID:", i, "Port:", tmp_dat$Port_Of_Landing[1])) +
    coord_cartesian(xlim=c(min(tmp_dat$LONGITUDE) - 0.1, max(tmp_dat$LONGITUDE) + 0.1), ylim=c(min(tmp_dat$LATITUDE) - 0.1, max(tmp_dat$LATITUDE) + 0.1))
  print(myplot)
}
```






















