---
title: "Returns to Port"
output: html_document
---

Document returns to port for Dungeness crab trips. 

This uses Blake's depth key to identify in port records.

Why? I'm thinking of truncating the VMS data, so that the input for home range / trip duration analysis only covers from when the vessel returned to port to sell fish, back to the last time the vessel left port. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "D:/VMS-repo") 
```
<br>

### Prep for running code

```{r}
rm(list=ls())
```
<br>
```{r include=FALSE}
library(adehabitatHR)
library(lubridate)
library(ggmap)
library(ggplot2)
library(dplyr)
library(maps)
library(rgdal)
library(rgeos)
library(raster)
```
<br>

Input / variables
```{r}
datfile <- "ProcessVMS/R_Output/interpolation/VMS_Outputs_wTARGET_10d_lookback_2010_final_cleaned_interpolated_regular.csv"
outfile <- "HomeRange/R_Output/2010_DCRB_TripCharacteristics_"
```
<br>


### Read in data
**VMS data**
```{r vms}
vms <- read.csv(datfile)
head(vms)
```

**Bathymetry**
```{r bathy}
bathy.grid <- readRDS("Input_Data/bathymetry/vms_composite_bath_spGrid.RDS")
```
<br>

### Find in-port points
Get bathymetry at VMS data points.
```{r add_depth}
vms_sp <- vms
coordinates(vms_sp) <- c("LONGITUDE", "LATITUDE") 
proj4string(vms_sp) <- CRS("+init=epsg:4326") # WGS 84
crs(bathy.grid) <- CRS("+init=epsg:4326") # WGS 84
bathy.points <- over(vms_sp, bathy.grid)$vms_composite_bath
vms <- mutate(vms, NGDC_M = bathy.points)
```
<br>

Use Blake's coding system to identify in-port points
```{r}
vms <- mutate(vms, in_port = ifelse(NGDC_M < -100000, "in_port", "offshore"))
```
<br>

Filter to Dungeness crab, negative bathymetry
```{r}
dcrb_vms <- vms %>%
  filter(TARGET_max=="DCRB") %>%
  filter(NGDC_M < 5)
```
<br>

### Check number of time vessel leaves port

Find trips where the vessel leaves port more than once (identified by the sequence of an in-port and then an offshore record).
```{r leaves}
leave_port <- dcrb_vms %>% arrange(Rec_ID, UTCDATETIM) %>%
  group_by(Rec_ID) %>%
  mutate(leave_port = in_port == "in_port" & lead(in_port == "offshore"), g=cumsum(leave_port)) %>%
  filter(leave_port | lag(leave_port)) %>%
  dplyr::select(Rec_ID, g) %>%
  distinct()
sum(leave_port$g > 1) / length(leave_port$g)
```
<br>

Now read in duration of each trip, in days
```{r}
duration <- read.csv("ProcessVMS/R_Output/WEAR_Workshop/2010_2016-2018_regularizedVMS_trip_duration.csv")
duration <- filter(duration, year == 2010)
```
<br>

Match to the leave port data frame, using Rec_ID
```{r}
leave_port <- left_join(leave_port, duration, by="Rec_ID")
```
<br>

```{r}
ggplot(leave_port, aes(x=g, y=n.days, group = g)) +
  geom_boxplot() +
  xlab("Leaves Port") +
  ylab("Days of Trip")
```
<br>

### Check number of times vessel enters port

Find trips where the vessel leaves port more than once (identified by the sequence of an in-port and then an offshore record).
```{r enters}
enter_port <- dcrb_vms %>% arrange(Rec_ID, UTCDATETIM) %>%
  group_by(Rec_ID) %>%
  mutate(enter_port = in_port == "offshore" & lead(in_port == "in_port"), g=cumsum(enter_port)) %>%
  filter(enter_port | lag(enter_port)) %>%
  group_by(Rec_ID) %>%
  summarise(g = max(g))
sum(enter_port$g ==0) / length(enter_port$g)
```
<br>

Match to the leave port data frame, using Rec_ID
```{r}
enter_port <- left_join(enter_port, duration, by="Rec_ID")
enter_port <- mutate(enter_port, size_cat = ifelse(FINAL_LENGTH >= 40, "Large", "Small"))
```
<br>

```{r}
ggplot(enter_port, aes(x=g)) + 
  geom_histogram() + theme_bw()
ggplot(data=filter(enter_port, !is.na(FINAL_LENGTH)), aes(y=FINAL_LENGTH, x=g)) + 
  geom_point() + theme_bw() +xlab("Enters Port") + ylab("Vessel Length (ft)")
ggplot(enter_port, aes(x=as.factor(g), y=n.days)) +
  geom_boxplot() +
  xlab("Enters Port") +
  ylab("Days of Trip")
```
<br>

WTF is going on when the vessel enters port more than 5 times?
```{r}
weird_trips <- filter(enter_port, g > 5)
weird_trips_vms <- filter(dcrb_vms, Rec_ID %in% weird_trips$Rec_ID)
length(unique(weird_trips$Rec_ID))
```
<br>


```{r plot}
data(stateMapEnv)
states_df <- map_data("state") %>%
  filter(region == "california" | region=="oregon" | region=="washington")
for(trip in unique(weird_trips$Rec_ID)){
  tmp_vms <- filter(weird_trips_vms, Rec_ID==trip)
  myplot<- ggplot() +
    geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
    geom_point(data=tmp_vms, aes(x=LONGITUDE, y=LATITUDE, col=westcoastdate_notime), alpha=0.5) +
    geom_path(data=tmp_vms, aes(x=LONGITUDE, y=LATITUDE), col="grey10", alpha=0.5) +
    coord_cartesian(xlim=c(min(tmp_vms$LONGITUDE - 0.25), max(tmp_vms$LONGITUDE + 0.25)), ylim=c(min(tmp_vms$LATITUDE -0.25), max(tmp_vms$LATITUDE +0.25))) +
    ggtitle(trip)
  print(myplot)
}
```
<br>

```{r}
dcrb_vms %>% filter(Rec_ID == trip) %>% dplyr::select(Rec_ID, westcoastdate, Port_Of_Landing, avg_speed_recalc, NGDC_M, in_port)
```


  