---
title: "Returns to Port"
output: html_document
---

Document returns to port for Dungeness crab trips. 

This uses Blake's depth key to identify in port records.

Why? I'm thinking of truncating the VMS data, so that the input for home range / trip duration analysis only covers from when the vessel returned to port to sell fish, back to the last time the vessel left port. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/Mary.Fisher/Documents/VMS-repo") 
```
<br>

### Prep for running code

```{r}
rm(list=ls())
```
<br>
```{r include=FALSE}
library(adehabitatHR)
library(lubridate)
library(ggmap)
library(ggplot2)
library(dplyr)
library(maps)
library(rgdal)
library(rgeos)
library(raster)
```
<br>

Input / variables
```{r}
datfile <- "ProcessVMS/R_Output/interpolation/VMS_Outputs_wTARGET_10d_lookback_2010_final_cleaned_interpolated_regular.csv"
outfile <- "HomeRange/R_Output/2010_DCRB_TripCharacteristics_"

# file with port coordinates, including directory
coords_file = "Input_Data/port_coords_fromBlake_edited.csv"

## calendar years
years <- c(2010)

## width of buffer circle (in meters). applies to all ports but those that require smaller buffers
r = 3000

## port codes for any ports that require reduced buffer sizes. Default is only COS
ports_lowbuffer <- c("COS")

## width of buffer circle (in meters) for COS
r2 = 1500

## cutoff value for speed (m/s) -- 1m/s ~ 2 knots
speed_cutoff <- 0.75
```
<br>


### Read in data
**VMS data**
```{r vms}
vms <- read.csv(datfile)
head(vms)
```

**Bathymetry**
```{r bathy}
bathy.grid <- readRDS("Input_Data/bathymetry/vms_composite_bath_spGrid.RDS")
```
<br>

**Port Locations**
```{r}
portdat <- read.csv(coords_file)
if(any(is.na(portdat$lat))==TRUE){
  message("WARNING: some ports are missing coordinates. ports were removed from data frame.")
  portdat <- filter(portdat, !is.na(lat))
}
port_coords <- portdat
port_coords_reg <- portdat %>% filter(!(port_code %in% ports_lowbuffer))
port_coords_sm <- portdat %>% filter(port_code %in% ports_lowbuffer)
```
<br>


### Prep the port coordinates
*Note - to assign a project, `proj4string` requires coordinates in xy (lng/lat) order*
```{r}
coordinates(port_coords) <- c("Lon", "Lat") 
proj4string(port_coords) <- CRS("+init=epsg:4326") # WGS 84
port_coords_geo <- spTransform(port_coords, CRSobj = CRS("+init=epsg:32610"))

coordinates(port_coords_reg) <- c("Lon", "Lat") 
proj4string(port_coords_reg) <- CRS("+init=epsg:4326") # WGS 84
port_coords_reg_geo <- spTransform(port_coords_reg, CRSobj = CRS("+init=epsg:32610"))

coordinates(port_coords_sm) <- c("Lon", "Lat") 
proj4string(port_coords_sm) <- CRS("+init=epsg:4326") # WGS 84
port_coords_sm_geo <- spTransform(port_coords_sm, CRSobj = CRS("+init=epsg:32610"))
```
<br>

### Create buffer
```{r buffer}
buffer_polygons_reg <- gBuffer(port_coords_reg_geo, byid=TRUE, width=r)
buffer_polygons_sm <- gBuffer(port_coords_sm_geo, byid=TRUE, width=r2)
buffer_polygons  <- rbind(buffer_polygons_reg, buffer_polygons_sm)
```
<br>

### Find in-port points
https://www.nceas.ucsb.edu/scicomp/usecases/point-in-polygon 

Assign projection to VMS points
```{r}
vms_sp <- vms
coordinates(vms_sp) <- c("LONGITUDE", "LATITUDE") 
proj4string(vms_sp) <- CRS("+init=epsg:4326") # WGS 84
vms_geo <- spTransform(vms_sp, CRSobj = CRS("+init=epsg:32610"))
```
<br>


Identify data points within buffer zones
```{r}
in.port <- over(vms_geo, buffer_polygons)$port_code
vms <- mutate(vms, in_port = in.port)
```
<br>


Mark "remove" if the record is in port *and* the avg speed is < 1 m/s.
```{r remove_flag}
vms <- mutate(vms, to_rm = ifelse(!is.na(in_port) & avg_speed_recalc < speed_cutoff, "remove", "retain"))
```
<br>

Filter to Dungeness crab
```{r}
dcrb_vms <- vms %>%
  filter(TARGET_max=="DCRB")
```
<br>

### Check number of time vessel leaves port

Find trips where the vessel leaves port more than once (identified by the sequence of an in-port and then an offshore record).
```{r leaves}
leave_port <- dcrb_vms %>% arrange(Rec_ID, UTCDATETIM) %>%
  group_by(Rec_ID) %>%
  mutate(leave_port = to_rm=="remove" & lead(to_rm=="retain"), g=cumsum(leave_port)) %>%
  filter(leave_port | lag(leave_port)) %>%
  dplyr::select(Rec_ID, g) %>%
  distinct()
sum(leave_port$g > 1) / length(leave_port$g)
```
<br>

Now read in duration of each trip, in days
```{r}
duration <- read.csv("ProcessVMS/R_Output/WEAR_Workshop/2010_2016-2018_regularizedVMS_trip_duration.csv")
duration <- filter(duration, year == 2010)
```
<br>

Match to the leave port data frame, using Rec_ID
```{r}
leave_port <- left_join(leave_port, duration, by="Rec_ID")
```
<br>

```{r}
ggplot(leave_port, aes(x=g, y=n.days, group = g)) +
  geom_boxplot() +
  xlab("Leaves Port") +
  ylab("Days of Trip")
```
<br>

### Check number of times vessel enters port

Find trips where the vessel leaves port more than once (identified by the sequence of an in-port and then an offshore record).
```{r enters2}
enter_port <- dcrb_vms %>% arrange(Rec_ID, UTCDATETIM) %>%
  group_by(Rec_ID) %>%
  mutate(enter_port = to_rm == "retain" & lead(to_rm == "remove"), g=cumsum(enter_port)) %>%
  filter(enter_port | lag(enter_port)) %>%
  group_by(Rec_ID) %>%
  summarise(g = max(g))
sum(enter_port$g ==0) / length(enter_port$g)
```
<br>

Match to the leave port data frame, using Rec_ID
```{r}
enter_port <- left_join(enter_port, duration, by="Rec_ID")
enter_port <- mutate(enter_port, size_cat = ifelse(FINAL_LENGTH >= 40, "Large", "Small"))
```
<br>

```{r}
ggplot(enter_port, aes(x=g)) + 
  geom_histogram() + theme_bw()
ggplot(data=filter(enter_port, !is.na(FINAL_LENGTH)), aes(y=FINAL_LENGTH, x=g)) + 
  geom_point() + theme_bw() +xlab("Enters Port") + ylab("Vessel Length (ft)")
ggplot(enter_port, aes(x=as.factor(g), y=n.days)) +
  geom_boxplot() +
  xlab("Enters Port") +
  ylab("Days of Trip")
```
<br>

WTF is going on when the vessel enters port more than 5 times?
```{r}
weird_trips <- filter(enter_port, g > 5)
weird_trips_vms <- filter(dcrb_vms, Rec_ID %in% weird_trips$Rec_ID)
length(unique(weird_trips$Rec_ID))
```
<br>


```{r plot}
data(stateMapEnv)
states_df <- map_data("state") %>%
  filter(region == "california" | region=="oregon" | region=="washington")
for(trip in unique(weird_trips$Rec_ID)){
  tmp_vms <- filter(weird_trips_vms, Rec_ID==trip)
  myplot<- ggplot() +
    geom_polygon(data=states_df, aes(x=long, y=lat, group=group), fill="grey47") +
    geom_point(data=tmp_vms, aes(x=LONGITUDE, y=LATITUDE, col=westcoastdate_notime), alpha=0.5) +
    geom_path(data=tmp_vms, aes(x=LONGITUDE, y=LATITUDE), col="grey10", alpha=0.5) +
    coord_cartesian(xlim=c(min(tmp_vms$LONGITUDE - 0.25), max(tmp_vms$LONGITUDE + 0.25)), ylim=c(min(tmp_vms$LATITUDE -0.25), max(tmp_vms$LATITUDE +0.25))) +
    ggtitle(trip)
  print(myplot)
}
```
<br>

```{r}
dcrb_vms %>% filter(Rec_ID == trip) %>% dplyr::select(Rec_ID, westcoastdate, Port_Of_Landing, avg_speed_recalc, in_port, to_rm)
```


### Compare different methods of identifying "returns to port"


**ArcGIS polygon-based returns**

Get bathymetry at VMS data points.
```{r add_depth}
vms_sp <- vms
coordinates(vms_sp) <- c("LONGITUDE", "LATITUDE") 
proj4string(vms_sp) <- CRS("+init=epsg:4326") # WGS 84
crs(bathy.grid) <- CRS("+init=epsg:4326") # WGS 84
bathy.points <- over(vms_sp, bathy.grid)$vms_composite_bath
vms <- mutate(vms, NGDC_M = bathy.points)
```
<br>

Use Blake's coding system to identify in-port points
```{r}
vms <- mutate(vms, in_port_bathy = ifelse(NGDC_M < -100000, "in_port", "offshore"))
```
<br>

Filter to Dungeness crab
```{r}
dcrb_vms <- vms %>%
  filter(TARGET_max=="DCRB") %>%
  filter(NGDC_M < 5)
```
<br>

```{r eval=FALSE}
weird_trips_vms <- filter(dcrb_vms, Rec_ID %in% weird_trips$Rec_ID)
View(weird_trips_vms %>% dplyr::select(Rec_ID, DOCNUM, UTCDATETIM, LATITUDE, LONGITUDE, avg_speed_recalc, in_port, to_rm, NGDC_M, in_port_bathy) %>% arrange(Rec_ID, UTCDATETIM))
```


Find trips where the vessel leaves port more than once, according to bathymetry
```{r enters}
enter_port_bathy <- dcrb_vms %>% arrange(Rec_ID, UTCDATETIM) %>%
  group_by(Rec_ID) %>%
  mutate(enter_port = in_port_bathy == "in_port" & lead(in_port_bathy == "offshore"), bathy.g=cumsum(enter_port)) %>%
  filter(enter_port | lag(enter_port)) %>%
  group_by(Rec_ID) %>%
  summarise(bathy.g = max(bathy.g))
sum(enter_port_bathy$bathy.g ==0) / length(enter_port_bathy$bathy.g)
```
```{r}
enter_port_compare <- left_join(enter_port, enter_port_bathy, by="Rec_ID")
enter_port_compare <- mutate(enter_port_compare, diff.g = g-bathy.g)
ggplot(enter_port_compare, aes(x=diff.g)) +
  geom_histogram() +
  xlab("Enters Port (Buffer Zone -- ArcGIS)")
```
ArcGIS consistently reports one extra entrance to port than buffer zones.
*This is because the speed cannot be calculated for the last record of each trip, so a "to remove" designation cannot be provided under the buffer zone filtering. However, the ArcGIS filtering can still designate the data point as in port.*

```{r eval=FALSE}
test_trips <- unique(filter(enter_port_compare, diff.g==-1)$Rec_ID)[1:5]
test_trips_vms <- filter(dcrb_vms, Rec_ID %in% test_trips)
View(test_trips_vms %>% dplyr::select(Rec_ID, DOCNUM, UTCDATETIM, LATITUDE, LONGITUDE, avg_speed_recalc, in_port, to_rm, NGDC_M, in_port_bathy) %>% arrange(Rec_ID, UTCDATETIM))
```
<br>







