---
title: "Time Gaps in Raw VMS Data"
output: html_document
---



<br>
```{r "setup", include=FALSE}
knitr::opts_knit$set(root.dir = "D:/VMS-repo") 
```
<br>
<br>

### Prep for running code
Clear workspace
```{r}
rm(list=ls())
```
<br>

Install packages
```{r include=FALSE}
library(dplyr)
library(tidyr)
library(lubridate) 
library(ggplot2)
library(move)
```
<br>

Create Objects / Data Frames
```{r}
## directory where filtered VMS data is stored as .csv file
VMS_dir = "ProcessVMS/R_Output/match/unfiltered/"

## calendar years
y <- 2016
```
<br>

### Read in data
```{r}
vms <- read.csv(paste0(VMS_dir,"VMS_Outputs_wTARGET_10d_lookback_", y, "_all.csv"))
```
<br>


#### Ensure that R recognises the timestamp as a POSIXct class
```{r}
vms$westcoastdate <- ymd_hms(vms$westcoastdate, tz= "America/Los_Angeles")
class(vms$westcoastdate)
vms$UTCDATETIM <- ymd_hms(vms$UTCDATETIM, tz= "UTC")

```
<br>


#### Deal with duplicates
Ensure that there are no duplicated dates/times under the same Rec_ID
```{r}
dim(vms)
vms_nodup <- vms[!duplicated(vms[,c("Rec_ID","UTCDATETIM")]),]
dim(vms_nodup)

cat("Proportion of VMS records removed for being duplicate records:", 1-dim(vms_nodup)[1]/dim(vms)[1])
```
<br>

#### Order timestamps within each individual
```{r}
vms_nodup <- vms_nodup %>%
  arrange(Rec_ID, UTCDATETIM)
```
<br>

### Create `move` object, with trip as "id"
```{r}
vms_nodup$DECLARATIO[vms_nodup$DECLARATIO == "N/A"] <- NA
```
<br>
```{r}
colnames(vms_nodup)
```
<br>
```{r}
start.time <- Sys.time()
movedat <- move(x=vms_nodup$LONGITUDE,
                y=vms_nodup$LATITUDE,
                time=vms_nodup$UTCDATETIM,
                data=vms_nodup,
                proj=sp::CRS("+init=epsg:4326"),
                animal=vms_nodup$Rec_ID)
Sys.time() - start.time
```
<br>

Use `timeLag` to get the amount of time that has elapsed between each successive VMS record.
```{r}
timestep <- timeLag(movedat, units="secs")
```
<br>
Add time lags back into move object
```{r}
movedat$timelag <- unlist(lapply(timestep, c, NA))
```
<br>

Create a data frame; thin it out
```{r}
vmsdat_move <- as.data.frame(movedat)
vmsdat_move <- vmsdat_move %>%
  dplyr::select(c(Port_Of_Landing, Rec_ID, LATITUDE, LONGITUDE, westcoastdate, timelag)) %>%
  arrange(Rec_ID, westcoastdate)
head(vmsdat_move)
```
<br>


Plot all time gaps
```{r}
ggplot(data=filter(vmsdat_move, !is.na(timelag)), aes(x=timelag/3600)) +
  geom_histogram() +
  xlab("Time Between VMS Records (hrs)")


ggplot(data=filter(vmsdat_move, !is.na(timelag) & timelag < 21600), aes(x=timelag/3600)) +
  geom_histogram() +
  xlab("Time Between VMS Records (hrs)")
```




