---
title: "Matching VMS Data to Bathymetry"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

10/30/19: Checking why there are so many VMS points without a matching 5km grid cell

```{r libs,message=FALSE,warning=FALSE}
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(raster)
library(rnaturalearth)
library(knitr)
library(ggsci)
library(extrafont)
library(marmap)

# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="Gill Sans MT",size=12,color="black"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)

```


## Import Files

```{r}
# load RData
fp <- "C:/Users/Owen.Liu/Documents/NWFSC Research/Dungeness Fleet Dynamics/VMS DATA CONFIDENTIAL/Processed Data/VMS/"
load(file=paste0(fp,"VMS_2009_2018_clean.RData"))

vms_sf<-readRDS(file=paste0(fp,"VMS_int_reg_w_grid.rds"))

# 5x5 grid shapefile
# grd <- read_sf(here::here("Input_Data","5x5 Grid","five_km_grid_polys_geo.shp"))
grd <- read_sf(here::here("Input_Data","5x5 Grid","regions_master_final_lamb.shp"))
names(grd)

# coastline
coaststates <- ne_states(country='United States of America',returnclass = 'sf') %>% 
  filter(name %in% c('California','Oregon','Washington')) %>% 
  st_transform(st_crs(vms_sf))

```

## Check Depth

One issue may be just that the VMS points that don't match are on land; if so, they should have positive values when matched to a bathymetry layer. Let's add depth.

```{r}
# Bathymetry layer
bathy <- readGEBCO.bathy(file="C:/Users/Owen.Liu/Documents/NWFSC Research/Dungeness Fleet Dynamics/GEBCO Grid/GEBCO_2019_-132.0_50.0_-115.0_29.0.nc")

# match depth (takes awhile to do the join)
vms_pts <- vms_sf %>% st_transform(4326) %>% st_coordinates()
depths <- vms_pts %>% 
    get.depth(bathy,.,locator=FALSE)
depths <- depths %>% as_tibble() %>% mutate(depth_m=depth*0.3048)

# join depth data
vms_sf <- vms_sf %>% bind_cols(dplyr::select(depths,lon,lat,depth_m))

# add binary indicator as to whether or not an observations has grid data
vms_sf <- vms_sf %>%
  mutate(has_grid=!is.na(GRID5KM_ID))

vms_sf %>% 
  ggplot(aes(depth_m,..density..,fill=has_grid))+
  geom_density(alpha=0.7)+
  coord_cartesian(xlim=c(-100,100))+
  geom_vline(xintercept=0)+
  labs(x="Depth (m)",title="Depth Distribution of VMS points")+
  scale_fill_npg(name="Has 5KM Grid")
```

So it seems like a lot of the points that weren't assigned a grid cell are likely on land (positive bathymetry). We can zoom in on a couple of places and see if this is true.

## Map Key Areas

```{r zoom}
# Monterey
monterey <- vms_sf %>% 
  filter(lat<37,lat>36.3,lon< -121.4,lon>-122.5)
bbox_monterey <- st_bbox(monterey)
monterey_points <- ggplot()+
  geom_sf(data=coaststates)+
  geom_sf(data=monterey,aes(col=has_grid,fill=has_grid),size=0.5,alpha=0.5)+
  scale_color_npg(name="Has 5km Grid\nassigned")+
  scale_fill_npg(name="Has 5km Grid\nassigned")+
  xlim(bbox_monterey[1],bbox_monterey[3])+
  ylim(bbox_monterey[2],bbox_monterey[4])+
  labs(title="Monterey VMS Points")

# Eureka
eureka <- vms_sf %>% 
  filter(lat<41.3,lat>40.5,lon< -123.5,lon>-124.5)
bbox_eureka <- st_bbox(eureka)
eureka_points <- ggplot()+
  geom_sf(data=coaststates)+
  geom_sf(data=eureka,aes(col=has_grid,fill=has_grid),size=0.5,alpha=0.5)+
  scale_color_npg(name="Has 5km Grid\nassigned")+
  scale_fill_npg(name="Has 5km Grid\nassigned")+
  xlim(bbox_eureka[1],bbox_eureka[3])+
  ylim(bbox_eureka[2],bbox_eureka[4])+
  labs(title="Eureka VMS Points")
monterey_points
eureka_points
```


