---
title: "Matching VMS Data to Spatial Grid"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here we match the interpolated VMS points from 2009 to a 5x5 grid provided by Blake Feist.

UPDATE 9/18/19: Using the topological union of multiple regions, not just the 5x5 grid. Because it is now a union, removed plotting of the output (because it's not as useful)

```{r libs,message=FALSE,warning=FALSE}
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(raster)
library(rnaturalearth)
library(knitr)
```


## Import Files

```{r}
# clean, interpolated vms data
vms <- read_csv("C:/Users/Owen.Liu/Documents/NWFSC Research/Dungeness Fleet Dynamics/VMS DATA CONFIDENTIAL/Processed Data/VMS/VMS_Outputs_wTARGET_10d_lookback_2009_final_cleaned_interpolated_regular.csv",col_types='dcTddccddTDdccccd')

# 5x5 grid shapefile
# grd <- read_sf(here::here("Input_Data","5x5 Grid","five_km_grid_polys_geo.shp"))
grd <- read_sf(here::here("Input_Data","5x5 Grid","regions_master_final_lamb.shp"))
names(grd)

meta <- readxl::read_xlsx(here::here("Input_Data","5x5 Grid","regions_master_final_lamb_metadata.xlsx")) 

meta %>% kable()

```

## Match Points

```{r}
# convert vms to spatial object (longitude/latitude)
vms_sf <- vms %>%
  st_as_sf(coords=c('LONGITUDE','LATITUDE'),crs=4326) %>% 
  # then, convert to planar projection to match the grid
  st_transform(st_crs(grd))

# do the join
pt <- proc.time()
vms_grd_match <- vms_sf %>%
  st_join(grd)
proc.time()-pt
```

## Check

Did the join work? How many unique 5km, 25km, and 50km cells are represented in the 2009 data?

```{r checks}
names(vms_grd_match)

length(unique(vms_grd_match$GRID5KM_ID))

length(unique(vms_grd_match$BAND_25KM))

length(unique(vms_grd_match$BAND_50KM))
```

## Plot [UPDATE 9.18.19: NOT RUN]

We can plot the output, with a raster representation of the grid for faster plotting

```{r, eval=FALSE}
# transform to a projection that uses meters instead of lat/long
grd_proj <- grd %>% st_transform(32610)
grd_rast <- fasterize(grd_proj,raster = raster(grd_proj,res=5000),field="GRID5KM_ID")
# set -99999 to NA
grd_rast[grd_rast==-99999] <- NA

vms_proj = vms_grd_match %>% st_transform(32610)
```

Plot points and raster grid

```{r, eval=FALSE}
# raster to dataframe for plotting
grd_rast_df <- rasterToPoints(grd_rast) %>% as_tibble() %>% rename(grid_id=layer)

# coastline for plotting
coast <- ne_states(country='United States of America',returnclass = 'sf') %>% 
  filter(name %in% c('California','Oregon','Washington')) %>% 
  st_transform(32610)

bbox_vms <- st_bbox(vms_proj)
ggplot()+
  geom_raster(data=grd_rast_df,aes(x,y,fill=grid_id))+
  geom_sf(data=vms_proj,size=0.5)+
  scale_fill_continuous(name="5 Km Grid ID")+
  lims(x=c(bbox_vms[1],bbox_vms[3]),y=c(bbox_vms[2],bbox_vms[4]))
```

## Write out

```{r write}
# write a non-spatial version of the joined data
vms_grd <- vms_grd_match
st_geometry(vms_grd) <- NULL

write_csv(vms_grd,"C:/Users/owenr/Documents/postdocs/NOAA NRC/NWFSC work/VMS DATA CONFIDENTIAL/Processed Data/VMS/VMS_Outputs_wTARGET_10d_lookback_2009_final_cleaned_interpolated_regular_w_5k_grid.csv")
```


## All Data

Check to see how fast (or slow) it is to match all of the VMS points from all years. For this we will use the combined (interpolated and regularized) data from `create_Rdata.Rmd`.

### Import

```{r import all}
# clear wrkspace
rm(list=setdiff(ls(),'grd'))

# load RData
fp <- "C:/Users/Owen.Liu/Documents/NWFSC Research/Dungeness Fleet Dynamics/VMS DATA CONFIDENTIAL/Processed Data/VMS/"
load(file=paste0(fp,"VMS_2009_2018_clean.RData"))
```

### Match

```{r}
# convert VMS to spatial object
pt <- proc.time()
vms_all_sf <- vms_int_reg %>%
  st_as_sf(coords=c('LONGITUDE','LATITUDE'),crs=4326) %>% 
  # then, convert to planar projection to match the grid
  st_transform(st_crs(grd))
proc.time()-pt
```

Took ~3m to do the conversion. Now for the join...

```{r join to grid}
# do the join
pt <- proc.time()
vms_all_grd_match <- vms_all_sf %>%
  st_join(grd)
proc.time()-pt
```

The join took ~15 minutes.

### Add BIAs (10/15/19)

We can add whale BIAs into the data as well. They have previously been matched to the 5km grid, courtesy of Sam Woodman.

```{r add BIAs}
BIA <- read_csv(here('Input_Data','Grid5km_BIA_overlap.csv'))
BIA %<>% select(GRID5KM_ID,BIA_mn,BIA_bm)

vms_all_grd_match %<>% left_join(BIA,by='GRID5KM_ID')
```

### Save and Write

```{r}
saveRDS(vms_all_grd_match,file=paste0(fp,"VMS_int_reg_w_grid.rds"))
```






