---
title: "Spatiotemporal Patterns in Vertical Lines"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r pkgs, include=FALSE,message=F,warning=FALSE}
library(raster)
library(tidyverse)
library(knitr)
library(extrafont)
library(ggsci)
library(lubridate)
library(sf)
library(geosphere)
library(magrittr)
library(rnaturalearth)
library(gstat)
library(data.table)
library(fasterize)
library(gganimate)

# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="Gill Sans MT",size=12,color="black"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)
```

# Introduction

The purpose of this document is to use the outputs from a vertical line/crab pot simulation estimation to investigate preliminary spatial and temporal patterns in the estimated number of traps in the water. We will do this by using a spatial grid to aggregate total crab traps across space and time.

```{r import data}
fp <- "C:/Users/Owen.Liu/Documents/NWFSC Research/Dungeness Fleet Dynamics/VMS DATA CONFIDENTIAL/Processed Data/VMS/"
fp_tix <- "C:/Users/Owen.Liu/Documents/NWFSC Research/Dungeness Fleet Dynamics/VMS DATA CONFIDENTIAL/Processed Data/Fish tickets/"

# fish tickets
load(file=paste0(fp_tix,"FishTix_all_2009_2018.RData"))

# a coastline, for plotting
# coastline for plotting
coaststates <- ne_states(country='United States of America',returnclass = 'sf') %>% 
  filter(name %in% c('California','Oregon','Washington','Nevada'))
coastline <- ne_coastline(scale='medium',returnclass = 'sf') %>% 
  st_crop(st_bbox(coaststates))

# Load if not already loaded
simulated_traps <- readRDS(file=paste0(fp,"sim_traps_speed_depth_filters.rds")) %>% ungroup()

# traps locations only (discard VMS data for now), attached to fish ticket data
traplocs <- simulated_traps %>% select(Rec_ID,traps) %>% 
  left_join(fishtix_matched_all,by="Rec_ID") %>% 
  select(Rec_ID,date,year,month,pacfin_port_code,port_group_code,agency_code,drvid,FINAL_LENGTH,traps) %>% 
  distinct(Rec_ID,.keep_all = T) %>% 
  # count number of traps associated with each trip
  mutate(ntraps=purrr::map_int(traps,nrow))

# Vessel monthly weighting factors
vessel_month_allweights<-readRDS(file=paste0(fp,"vessel_month_all_weights.rds"))

# 5x5 grid shapefile
# grd <- read_sf(here::here("Input_Data","5x5 Grid","five_km_grid_polys_geo.shp"))
grd <- read_sf(here::here("Input_Data","5x5 Grid","regions_master_final_lamb.shp"))
names(grd)

meta <- readxl::read_xlsx(here::here("Input_Data","5x5 Grid","regions_master_final_lamb_metadata.xlsx")) 

meta %>% kable()
```

```{r grid fxn,warning=FALSE,fig.height=8,fig.width=4}

# Function to build a grid from a given year/month's simulated traps data

# raster representation of grid.
# transform to a projection that uses meters instead of lat/long
grd_rast <- fasterize(grd,raster = raster(grd,res=5000),field="GRID5KM_ID")
# set -99999 to NA
grd_rast[grd_rast==-99999] <- NA

# add trap weights to traps list-column
vessel_month_allweights %>% ungroup() %>%  mutate(month=month.name[month]) %>% right_join(traplocs)->traplocs

traplocs %<>% mutate(traps=purrr::map2(traps,combined_weight,function(df,wt) mutate(df,wt=wt)))

count_traps <- function(yr,mth){
  mth_name <- month.name[mth]
  samp <- traplocs %>%
    ungroup() %>%
    filter(year==yr,month==mth_name)
    # mutate(traps_xy=purrr::map(traps,function(sfdf){st_coordinates(sfdf) %>% as_tibble(.name_repair = 'minimal') %>% set_names(c('lon','lat'))})) %>% 
    # select(traps_xy) %>% 
    # unnest()
  if(nrow(samp)==0){
    print(paste(mth_name,yr,"had zero observations."))
    zero_rast <- setValues(grd_rast,NA)
    names(zero_rast) <- paste0(mth_name,yr)
    return(zero_rast)
  } else{
    samp_sf<-setDT( samp )[, rbindlist( traps ) ] %>%
    sf::st_as_sf(crs=4326) %>% 
  # samp_sf <- samp %>%
  #   st_as_sf(coords=c('lon','lat'),crs=4326) %>% 
    # then, convert to planar projection to match the grid
    st_transform(st_crs(grd)) %>% 
    select(wt) %>% 
    as_Spatial()
  out_rast <- raster::rasterize(samp_sf,grd_rast,field='wt',fun='sum')/25
  names(out_rast) <- paste0(mth_name,yr)
  # out <- samp_sf %>%
  #   mutate(ind=row_number()) %>% 
  #   st_join(grd) %>% 
  #   ungroup() %>% 
  #   group_by(GRID5KM_ID) %>% 
  #   summarise(ntraps=n_distinct(ind),traps_per_sq_km=ntraps/25)
  # if(plot){
  #   st_geometry(out) <- NULL
  #   out_plot <- grd %>% 
  #     dplyr::select(GRID5KM_ID) %>% 
  #     left_join(out) %>% 
  #     filter(!is.na(traps_per_sq_km)) %>% 
  #     ggplot()+
  #     geom_sf(aes(fill=traps_per_sq_km),col=NA)+
  #     geom_sf(data=coaststates,col=NA)+
  #     scale_fill_viridis_c(name="Number of Traps\nper sq. km")+
  #     coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
  #     theme(axis.text.x.bottom = element_text(angle=45,vjust=+0.5,hjust=0))
    # return(out_plot)
  # } else return(out)
  print(paste(mth_name,yr,"done."))
  return(out_rast)
  }

}

# sample map
p <- proc.time()
dec2013 <- count_traps(2013,12)
proc.time()-p
dec2013
plot(dec2013)
# data.table version (dunno why i can't get something like this to work in tidy) 
#   traps_sf<-setDT( samp )[, rbindlist( traps ) ] %>% 
#     sf::st_as_sf()
```

Apply to all years/months' data.

```{r apply fxn}
p <- proc.time()
years_months <- crossing(yr=2009:2018,mth=1:12)
trap_density_all_years <- purrr::map2(years_months$yr,years_months$mth,count_traps)
proc.time()-p
trap_density_stack <- raster::stack(trap_density_all_years)
saveRDS(trap_density_stack,file=paste0(fp,"trap_density_all_weights.rds"))
```

# Animate output

We can then represent the output in an animation

```{r interpolate output,warning=FALSE,fig.height=8,fig.width=4}
# RasterStack to long form xy
all_obs_xy <- purrr::map(as.list(trap_density_stack),function(x){
  mth <- names(x)
  y<-rasterToPoints(x) %>% as_tibble() %>% set_names(c('x','y','density'))
  z <-y %>% mutate(month=mth)
  z
}) %>% bind_rows()

#bounding box for animation
bbox=c(825064,444947,1186239,2016996)

# test plot
coaststates %<>% st_transform(st_crs(grd))
test<-all_obs_xy %>% 
  filter(month=="December2009")
test %>% ggplot()+
    geom_tile(aes(x,y,fill=density),na.rm=T,alpha=0.8,interpolate=FALSE)+
    geom_sf(data=coaststates,col=NA,fill='gray50')+
    scale_fill_viridis(na.value=NA,option="C",name="Traps per\nsq. km",breaks=seq(0,200,by=50),limits=c(0,200))+
    coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
    labs(x='',y='',fill='Traps per\nsq. km',title='{current_frame}')+
    theme(axis.text.x.bottom = element_text(angle=45),
          legend.position = c(0.7,0.6),
          title=element_text(size=16))

# ordering factor for frames of the animation
all_obs_xy %<>% mutate(labeler=factor(month,levels=unique(month))) 
# produce the animation
library(viridis)
traps.gif <- all_obs_xy %>%
  filter(density>0) %>% 
  ggplot()+
    geom_tile(aes(x,y,fill=density),na.rm=T,alpha=0.8,interpolate=TRUE)+
    geom_sf(data=coaststates,col=NA,fill='gray50')+
    scale_fill_viridis(na.value=NA,option="C",name="Traps per\nsq. km",breaks=seq(0,200,by=50),limits=c(0,200))+
    coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
    labs(x='',y='',fill='Traps per\nsq. km',title='{current_frame}')+
    theme(axis.text.x.bottom = element_text(angle=45),
          legend.position = c(0.7,0.6),
          title=element_text(size=16))+
    transition_manual(labeler)
animate(traps.gif,fps=1,width=500,height=1000)

#save the animation
fp_plots <- "C:/Users/Owen.Liu/Documents/NWFSC Research/Dungeness Fleet Dynamics/VMS DATA CONFIDENTIAL/Processed Data/owen processed/plots/"
anim_save(filename=paste0(fp_plots,"trap_density_all_years_combined_weights.gif"))

# same map by log traps
# log.traps.gif <- all_obs_xy %>%
#   filter(density>0) %>% 
#   ggplot()+
#     geom_tile(aes(x,y,fill=log(density)),na.rm=T,alpha=0.8,interpolate=TRUE)+
#     geom_sf(data=coaststates,col=NA,fill='gray50')+
#     scale_fill_viridis(na.value=NA,option="C",name="Traps per\nsq. km")+
#     coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
#     labs(x='',y='',fill='Traps per\nsq. km',title='{current_frame}')+
#     theme(axis.text.x.bottom = element_text(angle=45),
#           legend.position = c(0.7,0.6),
#           title=element_text(size=16))+
#     transition_manual(labeler)
# animate(traps.gif,fps=1,width=500,height=1000)
# anim_save(filename=paste0(fp_plots,"log_trap_density_all_years.gif"))
```

```{r distribution of trap densities}
all_obs_xy %>% 
  filter(!is.na(density),density>0) %>% 
  ggplot(aes(density))+
  geom_density(fill='red',alpha=0.98,col=NA)+
  scale_x_continuous(limits=c(0,150))+
  labs(x="Density (Traps per sq. Km)",y="Kernel Density",title="Distribution of Positive Trap Densities\nall years and months")
```

```{r comparison to Blake}
# read in Blake's estimated data
blake<-readxl::read_xlsx(paste0(fp,"Monthly DCRB VMS pings per 5km grid cell 4Owen 101019.xlsx"))
grd_match <- tibble(GRID5KM_ID=getValues(grd_rast)) %>% left_join(blake,by="GRID5KM_ID")

# an example raster: for December, 2013
blake_rast <- raster(grd_rast)
values(blake_rast) <- grd_match$DUN2013_12
# reclassify zeroes to NA to match vertical line model
blake_rast <- reclassify(blake_rast,rcl=c(0,0,NA),right=NA)

test <- stack(blake_rast,trap_density_stack$December2013)
plot(test)
# do a correlation
cor(getValues(blake_rast),getValues(trap_density_stack$December2013),use="na.or.complete")

```

Correlation is not that high. Let's try with all the months in the data

```{r corr all months}
# Tibble to hold outputs of correlation
cor_table <- crossing(year=2009:2016,month=1:12,correlation=NA)
# Blake's analysis only goes through 06/2016
cor_table %<>% slice(-((nrow(cor_table)-5):nrow(cor_table)))

get_cor <- function(yr,mth){
  
  #parse correct blake year/month
  blake_var <- paste0("DUN",yr,"_",ifelse(mth>9,mth,paste0(0,mth)))
  # make raster and assign values
  blake_rast <- raster(grd_rast)
  values(blake_rast) <- grd_match[[blake_var]]
  # reclassify zeroes to NA to match vertical line model
  blake_rast <- reclassify(blake_rast,rcl=c(0,0,NA),right=NA)
  # grab correct owen year/month
  owen_var <- paste0(month.name[[mth]],yr)
  owen_rast <- trap_density_stack[[owen_var]]
  # do the correlation
  out <- cor(getValues(blake_rast),getValues(owen_rast),use="na.or.complete")
  # print(paste(month.name[mth],yr,"done."))
  out
}

cor_table %<>% mutate(correlation=map2_dbl(year,month,get_cor))

cor_table %>% 
  mutate(day=1) %>% 
  unite(date,year,month,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%m-%d",tz="America/Los_Angeles")) %>% 
  ggplot(aes(date,correlation))+
  geom_point()+
  geom_line()+
  labs(x="Date",y="Correlation",title="Correlation between vertical line model\nand Feist et al.")

head(arrange(cor_table,desc(correlation)))

## NOTE: ANOTHER OPTION?
#https://statnmap.com/2018-01-27-spatial-correlation-between-rasters/
```

## Aggregation at Larger Spatial Scales

Now that we have traps on the 5km grid scale, we can aggregate over larger spatial scales (counties, marine regions).

```{r}
subregion_grd <- grd %>% select(SUBREGION) %>% 
  group_by(SUBREGION) %>% 
  summarise()

ggplot(subregion_grd)+
  geom_sf(aes(fill=SUBREGION))
# for each month, convert raster to points and then aggregate
```

Let's make a plot for average March number of traps by these blocks.

IMPORTANT: The measurements in our `trap_density_stack` RasterStack are in traps per square kilometer. However, to sum the total number across aggregated blocks, we have to multiply these densities by 25 (1 trap per square kilometer equals 25 traps per 5kmx5km grid square). Then we can sum across 5x5 blocks.

```{r}
marchtraps <- subset(trap_density_stack,which(grepl("March",names(trap_density_stack)))) %>% 
  mean(na.rm=T)
# convert to counts from densities
marchtraps <- marchtraps*25

plot(marchtraps)

march_traps_points <- marchtraps %>% 
  rasterToPoints(spatial = TRUE) %>% 
  st_as_sf()

march_traps_regions <- subregion_grd %>% 
  st_join(march_traps_points) %>% 
  group_by(SUBREGION) %>% 
  summarise(tot_traps=sum(layer))

ggplot(march_traps_regions)+
  geom_sf(aes(fill=tot_traps))+
  labs(fill="Total Traps")

```

What about California macro blocks?

```{r}
ca_macro <- grd %>% select(CA_OFFSHOR) %>% 
  group_by(CA_OFFSHOR) %>% 
  summarise() %>% 
  filter(CA_OFFSHOR!=-999) %>% 
  mutate(mb=as.character(CA_OFFSHOR))

ggplot(ca_macro)+
  geom_sf(aes(fill=mb))

march_traps_mb <- ca_macro %>% 
  st_join(march_traps_points) %>% 
  group_by(mb) %>% 
  summarise(tot_traps=sum(layer)) %>% 
  filter(!is.na(tot_traps))

bbox_ca=st_bbox(march_traps_mb)

march_total_traps_plot <-ggplot(march_traps_mb)+
  geom_sf(aes(fill=tot_traps))+
  geom_sf_label(aes(label=round(tot_traps)))+
  geom_sf(data=coaststates,fill='gray50')+
  scale_fill_viridis(breaks=seq(0,40000,by=10000),labels=seq(0,40000,by=10000),limits=c(0,40000))+
  coord_sf(xlim=c(bbox_ca[1],bbox_ca[3]),ylim=c(bbox_ca[2],bbox_ca[4]))+
  labs(fill="Total Traps",title="Average Total Traps by Macroblock\nMarch, all years",x="",y="")+
  theme(legend.position = c(0.8,0.7),
        legend.background = element_rect(fill='white'))

march_total_traps_plot

ggsave(filename = paste0(fp_plots,"march_traps_CA_macroblocks.png"),march_total_traps_plot,h=6,w=6)
```

## Comparison to Sam Macks' thesis

We can compare our estimates of trap densities to those calculated by Sam Macks in her thesis.
For example, for 2013-2014 to 2016-17

```{r}
# Pull out data for November-February, 2013 to 2017
traps_macks_range <- all_obs_xy %>% 
  mutate(date=parse_date(month,format="%B%Y")) %>% 
  filter(date>=as.Date("2013-11-01"),date<=as.Date("2017-02-01")) %>% 
  mutate(month=month(date),year=year(date)) %>% 
  filter(month %in% c(11,12,1,2)) %>% 
  # average by month/year
  group_by(x,y,month) %>% 
  summarise(mean_density=mean(density,na.rm=T)) %>% 
  ungroup() %>% 
  mutate(monthname=factor(month.name[month],levels=c("November","December","January","February"))) %>% 
  filter(y<bbox_ca[4],y>bbox_ca[2])

ca_macks_comp_plot <- traps_macks_range %>% 
  ggplot()+
  geom_tile(aes(x,y,fill=mean_density),na.rm=T,alpha=0.8)+
  geom_sf(data=coaststates,col=NA,fill='gray50')+
  facet_wrap(~monthname,nrow=1)+
  scale_fill_viridis(na.value=NA,option="C",name="Traps per\nsq. km")+
  coord_sf(xlim=c(bbox_ca[1]+350000,bbox_ca[3]-100000),ylim=c(bbox_ca[2],bbox_ca[4]))+
  labs(x='',y='',fill='Traps per\nsq. km',title='Average Traps Density\nNovember to February, 2013-2017')+
  theme(axis.text.x.bottom = element_text(angle=45),
        title=element_text(size=16))
ca_macks_comp_plot

ggsave(filename = paste0(fp_plots,"CA_macks_comparison_plot.png"),ca_macks_comp_plot,h=6,w=10)
```



