---
title: "Interpolation"
output: html_document
---

Linear interpolation of time gaps that are either (1) less than 5 hours, or (2) accumulated when the vessel was in port. Uses the `move` package. 

<br>
```{r "setup", include=FALSE}
knitr::opts_knit$set(root.dir = "D:/VMS-repo") 
```
<br>
<br>

### Prep for running code
Clear workspace
```{r}
rm(list=ls())
```
<br>

Install packages
```{r include=FALSE}
library(dplyr)
library(tidyr)
library(lubridate) 
library(ggplot2)
library(move)
library(rgdal)
```
<br>

Create Objects / Data Frames
```{r}
## directory where filtered VMS data is stored as .csv file
VMS_dir = "C:/Users/owenr/Documents/postdocs/NOAA NRC/NWFSC work/VMS DATA CONFIDENTIAL/Processed Data/owen processed/clean/"

## calendar years
years <- c(2009)

## recalculate the average speed?
recalc_speed = TRUE
```
<br>



### Read in the data

First, the full vms data set
```{r rawdat}
for(i in seq(1,length(years))){
  y <- years[i]
  if(i == 1){
    vms <- read.csv(paste0(VMS_dir,"VMS_Outputs_wTARGET_10d_lookback_", y, "_final_cleaned.csv"))
  } else{
    tmp_vms <-read.csv(paste0(VMS_dir,"VMS_Outputs_wTARGET_10d_lookback_", y, "_final_cleaned.csv"))
    vms <- rbind(vms, tmp_vms)
  }
}

if("X" %in% colnames(vms)){
  vms <- dplyr::select(vms, -X)
}
cat("VMS data consists of ", length(unique(vms$Rec_ID)), " fishing trips.\n")
head(vms)
```
<br>

### Interpolate using a `move` object / `interpolateTime`
#### Ensure that R recognises the timestamp as a POSIXct class
```{r}
vms$westcoastdate <- ymd_hms(vms$westcoastdate, tz= "America/Los_Angeles")
class(vms$westcoastdate)
vms$UTCDATETIM <- ymd_hms(vms$UTCDATETIM, tz= "UTC")
vms$DECLARATIO[vms$DECLARATIO == "N/A"] <- NA
```
<br>

#### Order timestamps within each individual
```{r}
vms <- vms %>%
  arrange(Rec_ID, UTCDATETIM)
```
<br>

```{r move}
start.time <- Sys.time()
movedat <- move(x=vms$X_COORD,
                y=vms$Y_COORD,
                time=vms$UTCDATETIM,
                data=vms,
                proj=sp::CRS("+proj=utm +north +zone=10 +ellps=WGS84"),
                animal=vms$Rec_ID)
Sys.time() - start.time
```
<br>

#### Function to loop over each trip in a data frame and interpolate

Starting from a MoveStack object
```{r}
interpolate_trips <- function(movedata, timestep=60, units="mins", method = "euclidean", clean=TRUE){
  # number of trips
  n_trips <- n.indiv(movedata)
  # for loop
  for(i in seq(1,n_trips)){
    tmp_movedat <- movedata[[i]]
    inter_movedat <- interpolateTime(x=tmp_movedat, time=as.difftime(timestep, units=units), spaceMethod=method)
    inter_df <- as.data.frame(inter_movedat)
    if(clean){
      #clean up the data frame
      inter_df <- inter_df %>%
        dplyr::select(Rec_ID, DOCNUM, UTCDATETIM, X_COORD, Y_COORD, in_port, to_rm, timestamps, coords.x1, coords.x2, sensor)
    }
    
    if(i == 1){
      interpolated_data <- inter_df
    } else{
      interpolated_data <- rbind(interpolated_data, inter_df)
    }
    if(i%%500==0){
      message("completed interpolation for ", i, " out of ", n_trips, " trips.")
    }
  }
  return(interpolated_data)
}
```
<br>

### Linear Interpolation

```{r interpolate}
start.time <- Sys.time()
interpolated_vms <- interpolate_trips(movedata = movedat, timestep=60, units="mins", method = "euclidean", clean=TRUE)
Sys.time() - start.time
```
On 2016 data, time difference of 1.357018 hours

<br>


### Regularized data: Integrated interpolated data into original data frame
The following will only run properly if the `clean` argument in the `interpolated_vms` function is set to TRUE.

First, separate interpolated data from the original data points so that new vms columns can be renamed
```{r}
orig_vms <- interpolated_vms %>%
  filter(sensor=="unknown") %>%
  dplyr::select(Rec_ID, DOCNUM, UTCDATETIM, X_COORD, Y_COORD, sensor)
new_vms <- interpolated_vms %>%
  filter(sensor=="interpolateTime") %>%
  dplyr::select(-UTCDATETIM, -X_COORD, -Y_COORD)
colnames(new_vms)[which(colnames(new_vms) == "timestamps")] <- "UTCDATETIM"
colnames(new_vms)[which(colnames(new_vms) == "coords.x1")] <- "X_COORD"
colnames(new_vms)[which(colnames(new_vms) == "coords.x2")] <- "Y_COORD"
head(new_vms)
```
<br>

Get trip / record-specific info for the original vms data points that were retained.
```{r}
orig_vms <- left_join(orig_vms, vms, by=c("Rec_ID", "DOCNUM", "UTCDATETIM", "X_COORD", "Y_COORD"))
```
<br>

Get trip info for the new vms data points. This originally pulled from the full `vms` data frame, but for a few trips, one of the vessel had multiple listed names. This duplicated time stamps when joining. This is avoided by pulling trip info from the `orig_vms` data frame.
```{r}
vms_trip_info <- orig_vms %>%
  dplyr::select(Rec_ID, DOCNUM, VESSEL_NAM, DECLARATIO, TARGET, Port_Of_Landing, TARGET_max) %>%
  distinct()
# one vessel had multiple names listed for trip id/ docnum
if(dim(vms_trip_info[duplicated(vms_trip_info[,c("Rec_ID", "DOCNUM")]),])[1] > 0){
  print(vms_trip_info[duplicated(vms_trip_info[,c("Rec_ID", "DOCNUM")]),])
  vms_trip_info <- vms_trip_info[!duplicated(vms_trip_info[,c("Rec_ID", "DOCNUM")]),]
}
new_vms <- left_join(new_vms, vms_trip_info, by=c("Rec_ID", "DOCNUM"))
sum(is.na(new_vms$TARGET))
```
<br>

Create calculable missing columns: west coast dates, latitude, longitude. 
```{r}
# west coast dates
new_vms <- new_vms %>%
  mutate(westcoastdate = with_tz(UTCDATETIM, tzone = "America/Los_Angeles"),
         westcoastdate_notime = as_date(westcoastdate))

# convert UTM x/y coordinates to lat/lon
new_vms_coords <- new_vms
coordinates(new_vms_coords) <- c("X_COORD", "Y_COORD")
new_vms.spatial <- SpatialPoints(new_vms_coords, proj4string=CRS("+proj=utm +north +zone=10 +ellps=WGS84"))
data.LL <-spTransform(new_vms.spatial, CRS("+proj=longlat +datum=WGS84"))
data.LL.df <- as.data.frame(data.LL); colnames(data.LL.df) <- c("LONGITUDE", "LATITUDE")

# add the new lat/lon into the dataset
new_vms <- new_vms %>%
  mutate(LONGITUDE = data.LL.df$LONGITUDE,
         LATITUDE = data.LL.df$LATITUDE)
head(new_vms)
```
<br>

Add in VMS Record Numbers for new VMS points
```{r}
new_recnos <- seq(from=(max(vms$VMS_RECNO)+10), length.out=length(new_vms$Rec_ID))
new_vms <- mutate(new_vms, VMS_RECNO = new_recnos)
```
<br>

Recombine
```{r}
orig_vms$westcoastdate_notime <- ymd(orig_vms$westcoastdate_notime)
vms_regular <- full_join(orig_vms, new_vms, by=colnames(new_vms))
# check record numbers
length(unique(vms_regular$VMS_RECNO))/length(vms_regular$VMS_RECNO)
```
<br>

Optional: Recalculate the average speed for each record, and save it as an extra column. **untested**
```{r}
if(recalc_speed){
  start.time <- Sys.time()
  # arrange data
  vms_regular <- vms_regular %>% arrange(Rec_ID, UTCDATETIM)
  # check for duplicates
  
  # convert to movestack
  movedat <- move(x=vms_regular$X_COORD,
                  y=vms_regular$Y_COORD,
                  time=vms_regular$UTCDATETIM,
                  data=vms_regular,
                  proj=sp::CRS("+proj=utm +north +zone=10 +ellps=WGS84"),
                  animal=vms_regular$Rec_ID)
  message("created MoveStack object")
  # calculate time and distance between records
  timestep <- timeLag(movedat, units="secs")
  diststep <- distance(movedat)
  message("calculated distance and time steps.")
  movedat$dt <- unlist(lapply(timestep, c, NA))
  movedat$dx <- unlist(lapply(diststep , c, NA))
  # turn back into data frame
  vms_reg_move <- as.data.frame(movedat)
  vms_reg_move <- vms_reg_move %>%
    dplyr::select(c(Rec_ID, X_COORD, Y_COORD, UTCDATETIM, dt, dx)) %>%
    arrange(Rec_ID, UTCDATETIM)
  # calculate new average speed
  vms_reg_move <- vms_reg_move %>%
    mutate(avg_speed_recalc = dx/dt)
  # join average speeds back to data frame
  vms_reg_move$Rec_ID  <- as.character(vms_reg_move$Rec_ID)
  vms_regular$Rec_ID <- as.character(vms_regular$Rec_ID)
  vms_regular <- left_join(vms_regular, vms_reg_move, by=c("Rec_ID", "UTCDATETIM", "X_COORD", "Y_COORD"))
  message("added average speeds to data frame.")
  Sys.time() - start.time
}
```
<br>


Reorder columns
```{r}
if(recalc_speed){
  vms_regular <- vms_regular[,c("Rec_ID","DOCNUM","UTCDATETIM","LATITUDE","LONGITUDE","VESSEL_NAM","DECLARATIO","X_COORD","Y_COORD","westcoastdate","westcoastdate_notime","avg_speed_recalc","TARGET","Port_Of_Landing","TARGET_max","sensor","VMS_RECNO")]
} else{
  vms_regular <- vms_regular[,c("Rec_ID","DOCNUM","UTCDATETIM","LATITUDE","LONGITUDE","VESSEL_NAM","DECLARATIO","X_COORD","Y_COORD","westcoastdate","westcoastdate_notime","TARGET","Port_Of_Landing","TARGET_max","sensor","VMS_RECNO")]
}
```
<br>

Write out
```{r}
write.csv(vms_regular, paste0(VMS_dir,"VMS_Outputs_wTARGET_10d_lookback_", y, "_final_cleaned_interpolated_regular.csv"), row.names=FALSE)
```
<br>



### All data: Integrated interpolated data into original data frame
The following will only run properly if the `clean` argument in the `interpolated_vms` function is set to TRUE.


Join new vms to original vms
```{r}
vms$westcoastdate_notime <- ymd(vms$westcoastdate_notime)
vms_all <- full_join(vms, new_vms, by=colnames(new_vms)[which(colnames(new_vms) != "sensor")])
# check record numbers
length(unique(vms_all$VMS_RECNO))/length(vms_all$VMS_RECNO)
```
<br>

For all NAs in sensor, write "data"
```{r}
vms_all$sensor <- as.character(vms_all$sensor)
vms_all$sensor[which(is.na(vms_all$sensor))] <- "data"
```
<br>

Optional: Recalculate the average speed for each record, and save it as an extra column. **untested**
```{r}
if(recalc_speed){
  start.time <- Sys.time()
  # arrange data
  vms_all <- vms_all %>% arrange(Rec_ID, UTCDATETIM)
  # convert to movestack
  movedat <- move(x=vms_all$X_COORD,
                  y=vms_all$Y_COORD,
                  time=vms_all$UTCDATETIM,
                  data=vms_all,
                  proj=sp::CRS("+proj=utm +north +zone=10 +ellps=WGS84"),
                  animal=vms_all$Rec_ID)
  message("created MoveStack object")
  # calculate time and distance between records
  timestep <- timeLag(movedat, units="secs")
  diststep <- distance(movedat)
  movedat$dt <- unlist(lapply(timestep, c, NA))
  movedat$dx <- unlist(lapply(diststep , c, NA))
  message("calculated distance and time steps.")
  # turn back into data frame
  vms_all_move <- as.data.frame(movedat)
  vms_all_move <- vms_all_move %>%
    dplyr::select(c(Rec_ID, X_COORD, Y_COORD, UTCDATETIM, dt, dx)) %>%
    arrange(Rec_ID, UTCDATETIM)
  # calculate new average speed
  vms_all_move <- vms_all_move %>%
    mutate(avg_speed_recalc = dx/dt)
  # join average speeds back to data frame
  vms_all_move$Rec_ID  <- as.character(vms_all_move$Rec_ID)
  vms_all$Rec_ID <- as.character(vms_all$Rec_ID)
  vms_all <- left_join(vms_all, vms_all_move, by=c("Rec_ID", "UTCDATETIM", "X_COORD", "Y_COORD"))
    message("added average speeds to data frame.")
  Sys.time() - start.time
}
```
<br>


Write out as csv
```{r}
write.csv(vms_all, paste0(VMS_dir,"VMS_Outputs_wTARGET_10d_lookback_", y, "_final_cleaned_interpolated_all.csv"), row.names=FALSE)
```
<br>

________________

### Alternative interpolation: starting from a data frame
Starting from a data frame, rather than a `moveStack` object, took about 30% longer to run. 

Function
```{r eval=FALSE}
interpolate_trips <- function(mydata, timestep=60, units="mins", method = "euclidean", clean=TRUE){
  # number of trips
  trip_ids <- unique(mydata$Rec_ID)
  n_trips <- length(trip_ids)
  # for loop
  for(i in seq(1,n_trips)){
    tmp_vms <- filter(mydata, Rec_ID == trip_ids[i])
    movedat <- move(x=tmp_vms$X_COORD,
                    y=tmp_vms$Y_COORD,
                    time=tmp_vms$UTCDATETIM,
                    data=tmp_vms,
                    proj=sp::CRS("+init=epsg:32610"),
                    animal=tmp_vms$Rec_ID)
    inter_movedat <- interpolateTime(x=movedat, time=as.difftime(timestep, units=units), spaceMethod=method)
    inter_df <- as.data.frame(inter_movedat)
    if(clean){
      #clean up the data frame
      inter_df <- inter_df %>%
        dplyr::select(Rec_ID, DOCNUM, UTCDATETIM, X_COORD, Y_COORD, timestamps, coords.x1, coords.x2, sensor)
    }
    
    if(i == 1){
      interpolated_data <- inter_df
    } else{
      interpolated_data <- rbind(interpolated_data, inter_df)
    }
    if(i%%500==0){
      message("completed interpolation for ", i, " out of ", n_trips, " trips.")
    }
  }
  return(interpolated_data)
}
```
Calls on function above.
```{r eval=FALSE}
start.time <- Sys.time()
interpolated_vms <- interpolate_trips(mydata=vms, timestep=60, units="mins", method = "euclidean", clean=TRUE)
Sys.time() - start.time
```
time difference: 1.792642 hrs






