---
title: "Filter for Time Gaps in VMS Data"
output: html_document
---

Last step before interpolation. Filters for time gaps over which it is unlikely that linear interpolation will be representative. According to the following rules:

IF the maximum time gap is > 4 hours OR composes > 25% of the trip, Remove the trip. UNLESS...

1. The maximum time gap occurs when the vessel is in port. <- Leave the trip
<br>
2. The maximum time gap is the second to last record. <- Write out the trip to another file for further examination.


<br>
```{r "setup", include=FALSE}
knitr::opts_knit$set(root.dir = "D:/VMS-repo") 
```
<br>
<br>

### Prep for running code
Clear workspace
```{r}
rm(list=ls())
```
<br>

Install packages
```{r include=FALSE}
library(dplyr)
library(tidyr)
library(lubridate) 
library(ggplot2)
library(move)
```
<br>

Create Objects / Data Frames
```{r}
## directory where filtered VMS data is stored as .csv file
VMS_dir ="C:/Users/owenr/Documents/postdocs/NOAA NRC/NWFSC work/VMS DATA CONFIDENTIAL/Processed Data/owen processed/clean/"

## calendar years
years <- c(2009)
```
<br>



### Read in the data

First, the full vms data set
```{r rawdat}
for(i in seq(1,length(years))){
  y <- years[i]
  if(i == 1){
    vms <- read.csv(paste0(VMS_dir,"VMS_Outputs_wTARGET_10d_lookback_", y, "_speed_cleaned_offshore.csv"))
  } else{
    tmp_vms <-read.csv(paste0(VMS_dir,"VMS_Outputs_wTARGET_10d_lookback_", y, "_speed_cleaned_offshore.csv"))
    vms <- rbind(vms, tmp_vms)
  }
}

if("X" %in% colnames(vms)){
  vms <- dplyr::select(vms, -X)
}
dim(vms)
head(vms)
```
<br>


### Calculate dt between records
#### Ensure that R recognises the timestamp as a POSIXct class
```{r}
vms$westcoastdate <- ymd_hms(vms$westcoastdate, tz= "America/Los_Angeles")
class(vms$westcoastdate)
vms$UTCDATETIM <- ymd_hms(vms$UTCDATETIM, tz= "UTC")
vms$DECLARATIO[vms$DECLARATIO == "N/A"] <- NA
```
<br>

#### Order timestamps within each individual
```{r}
vms <- vms %>%
  arrange(Rec_ID, UTCDATETIM)
```
<br>

#### Create `move` object, with trip as "id"
```{r}
start.time <- Sys.time()
movedat <- move(x=vms$LONGITUDE,
                y=vms$LATITUDE,
                time=vms$UTCDATETIM,
                data=vms,
                proj=sp::CRS("+init=epsg:4326"),
                animal=vms$Rec_ID)
Sys.time() - start.time
```
<br>

Use `timeLag` to get the amount of time that has elapsed between each successive VMS record.
```{r}
timestep <- timeLag(movedat, units="secs")
```
<br>

Add time lags back into move object
```{r}
movedat$timelag <- unlist(lapply(timestep, c, NA))
```
<br>

Create a data frame; thin it out
```{r}
vmsdat_move <- as.data.frame(movedat)
vmsdat_move <- vmsdat_move %>%
  dplyr::select(c(Port_Of_Landing, Rec_ID, LATITUDE, LONGITUDE, westcoastdate, avg_speed_ms, in_port, timelag)) %>%
  arrange(Rec_ID, westcoastdate)
head(vmsdat_move)
```
<br>

### Flag trips for removal

Number records in order for each trip
```{r}
vmsdat_move <- vmsdat_move %>%
  group_by(Rec_ID) %>%
  mutate(trip_dt = sum(timelag, na.rm=TRUE), recnum = seq(1, length(LATITUDE)), max_recnum = max(recnum)) %>%
  ungroup()
```
<br>

Grab the record with the greatest time gap between records
```{r}
vmsdat_maxDT <- vmsdat_move %>%
  group_by(Rec_ID) %>%
  top_n(1, wt=timelag)
```
<br>

Flag the trip if the timelag > 4 hours OR more than 25% total trip time
```{r}
vmsdat_maxDT <- vmsdat_maxDT %>%
  mutate(remove_trip = ifelse(timelag > (4*3600) | (timelag/trip_dt) > 0.40, "remove", "retain"))

n_remove = sum(vmsdat_maxDT$remove_trip == "remove")
cat("Flagged ", n_remove/length(vmsdat_maxDT$remove_trip)*100, "% (", n_remove,") trips for removal.")

cat(sum(vmsdat_maxDT$remove_trip == "remove" & !is.na(vmsdat_maxDT$in_port))/n_remove*100, "% removed trips were in-port timelags.")
cat(sum(vmsdat_maxDT$remove_trip == "remove" & (vmsdat_maxDT$max_recnum-vmsdat_maxDT$recnum == 1))/n_remove*100, "% removed trips were the second to last record in the trip.")
```
<br>

### Filter and write out data
Trips to retain
```{r}
# trips to retain
trips_retained <- vmsdat_maxDT %>%
  filter(remove_trip == "retain" | remove_trip == "remove" & !is.na(in_port)) %>%
  dplyr::select(Rec_ID)
# get the VMS records for those trips
vms_out <- vms %>%
  dplyr::filter(Rec_ID %in% trips_retained$Rec_ID) %>%
  dplyr::select(-dt,-dx)
# write out
for(y in years){
  tmp_out <- filter(vms_out, year(westcoastdate) == y)
  write.csv(tmp_out, paste0(VMS_dir,"VMS_Outputs_wTARGET_10d_lookback_", y, "_final_cleaned.csv"), row.names=FALSE)
}
```
<br>

Trips to remove
```{r}
# trips to retain
trips_removed <- vmsdat_maxDT %>%
  filter(!(Rec_ID %in% trips_retained$Rec_ID) & ((max_recnum-recnum) != 1)) %>%
  dplyr::select(Rec_ID)
# get the VMS records for those trips
vms_removed <- vms %>%
  filter(Rec_ID %in% trips_removed$Rec_ID) %>%
  dplyr::select(-dt)
# write out
for(y in years){
  tmp_out <- filter(vms_removed, year(westcoastdate) == y)
  write.csv(tmp_out, paste0(VMS_dir,"VMS_Outputs_wTARGET_10d_lookback_", y, "_timegaps_removed.csv"), row.names=FALSE)
}
```
<br>

Trips marked to remove, but for the second to last record in the trip
```{r}
# trips to remove
trips_removed_tmp <- vmsdat_maxDT %>%
  filter(!(Rec_ID %in% trips_retained$Rec_ID) & ((max_recnum-recnum) == 1)) %>%
  dplyr::select(Rec_ID)
# get the VMS records for those trips
vms_removed_tmp <- vms %>%
  filter(Rec_ID %in% trips_removed_tmp$Rec_ID) %>%
  dplyr::select(-dt)
# write out
for(y in years){
  tmp_out <- filter(vms_removed_tmp, year(westcoastdate) == y)
  write.csv(tmp_out, paste0(VMS_dir,"VMS_Outputs_wTARGET_10d_lookback_", y, "_timegaps_removed_temp.csv"), row.names=FALSE)
}
```
<br>









